### 2021-04-13 doc/introduction.Rmd
adding the builded vignettes to the repo

---
title: "Spatial Fuzzy CMean with R"
author: "Jeremy Gelb"
date: "02/04/2021"
output:
  rmarkdown::html_vignette:
    fig_width: 5
    fig_height: 5
    toc: true
    toc_depth: 2
    df_print: "tibble"
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: geocmeans_references.bibtex
link-citations: yes
colorlinks: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document is a short introduction to the package **geocmeans**. It implements a fuzzy classification method bringing spatial information and neighbouring in its calculation

@cai2007fast described in their article the method, originally applied in the analysis of brain imagery. The generalized version of the spatial fuzzy c-means is presented by @zhao2013kernel.

@gelb2021apport applied the method to socio-residential and environmental data and compared the results with oter unsupervised classification algorithms (in french).

There are actually numerous packages and functions to perform unsupervised classification in R (*hclust*, *kmeans*, *cmeans*, *factoextra*, etc.). However theses methods are not always well-suited to analyze spatial data. Indeed, they do not take into account spatial information like proximity or contiguity between observations. This may lead to solutions for which close observations end up in different groups despite the fact that they are very similar.

The package **ClustGeo** is to our knowledge the only package proposing an unsupervised classification method considering directly spatial proximity between observations. The proposed approach is very appealing because the user has to select a parameter (*alpha*) that controls the weight of the spatial distance matrix (calculated between observations with their locations) VS the semantic distance matrix (calculated  between observations with their variables).

However, this method belongs to the category of "hard-clustering" algorithms. Each observation ends up in one cluster/group. The main draw-back here is the difficulty to identify observations that are undecided, at the frontier of two clusters/groups. The soft or fuzzy clustering algorithms give more information because they calculate for each observation its "probability" of belonging to each group.

The algorithms SFCM (spatial fuzzy c-means) and SGFCM (spatial generalized fuzzy c-means) propose to combine the best of both worlds.

The package **geocmeans** is an implementation in R of these methods (originally developed for analysis of brain imagery). It comes with a set of functions to facilitate the analysis of the final membership matrices :

* calculating many quality indices (coming mainly from the package *fclust*)
* mapping the results
* giving summary statistics for each cluster/group.

## Loading the packages and the data

**geocmeans** comes with a toy dataset *LyonIris*, combining many demographic and environmental variables aggregated at the scale of the Iris (aggregated units for statistical information) in Lyon (France).

Before starting the analysis, it is required to standardize the data because most of the calculus is based on Euclidean distance (we plan to also include Manhattan distance in a future release).

```{r message=FALSE, warning=FALSE}
#charging packages and data
library(geocmeans)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(viridis)
library(spdep)

data(LyonIris)

#selecting the columns for the analysis
AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14",
                   "Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")

#rescaling the columns
Data <- LyonIris@data[AnalysisFields]
for (Col in names(Data)){
  Data[[Col]] <- scale(Data[[Col]])
}

#preparing some elements for further mapping
LyonIris$OID <- as.character(1:nrow(LyonIris))
FortiData <- broom::tidy(LyonIris,region="OID")
```

```{r include=FALSE}
#loading the pre-calculated results
load(system.file("extdata", "results_vignette_intro.rda",
                           package = "geocmeans", mustWork = TRUE))
```

## Classical Kmeans

To explore the dataset and choose a number of cluster/groups (k) we propose to start with a classical kmeans.

```{r warning=FALSE, fig.cap = "Impact of the number of groups on the explained variance", out.width = "50%", fig.pos="H", fig.align="center"}
#finding the best k by using the r2 of the classification
#trying for k from 2 to 10
R2s <- sapply(2:10,function(k){
  Clust <- kmeans(Data,centers=k,iter.max = 150)
  R2 <- Clust$betweenss / Clust$totss
  return(R2)
})

Df <- data.frame(K=2:10,
                 R2 = R2s)

ggplot(Df)+
  geom_line(aes(x=K,y=R2s))+
  geom_point(aes(x=K,y=R2s),color="red")+
  xlab("Number of groups")+
  ylab("R2 of classification")
```

By plotting the R-squared of the kmeans classification for each k between 2 and 10, we can see a first elbow at k=3. But this small number of groups leads to a classification explaining only 43% of the original data variance. We decide to keep k=4 to have one more group in the analysis.

Let us map the obtained groups.

```{r warning=FALSE, fig.cap="Clusters identified by classical k-means", fig.pos="H", fig.align="center"}
KMeanClust <-  kmeans(Data,centers=4,iter.max = 150)
LyonIris$Cluster <-paste("cluster",KMeanClust$cluster,sep="_")

#mapping the groups
DFmapping <- merge(FortiData,LyonIris,by.x="id",by.y="OID")

ggplot(data=DFmapping)+
  geom_polygon(aes(x=long,y=lat,group=group,fill=Cluster),color=rgb(0,0,0,0))+
  coord_fixed(ratio = 1)+
  scale_fill_manual(name="Cluster",values = c("cluster_1"="palegreen3",
                                 "cluster_2"="firebrick",
                                 "cluster_3"="lightyellow2",
                                 "cluster_4"="steelblue"))+
  theme( axis.title = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.ticks = ggplot2::element_blank()
    )

```

We can clearly distinguish 4 strong spatial structures, but with some mixing between the clusters.

We could now compare this solution with a classical c-means algorithm.

## Classical c-means and generalized c-means

The classical c-means is a simple method to perform fuzzy unsupervised classification. The package **geocmeans** proposes the function *CMeans*. We set the fuzzyness degree (m) to 1.5.

We do not present here the algorithm, but only the two main formulas.

The first one is used to update the values of the membership matrix at each iteration $u_ik$

$$u_{ik} = \frac{(||x_{k} - v{_i}||^{2}) ^{(-1/(m-1))}}{\sum_{j=1}^c(||x_{k} - v{_j}||^2 )^{(-1/(m-1))}}$$
And the second one to update the centers of the clusters

$$v_{i} = \frac{\sum_{k=1}^N u_{ik}^m(x_{k})}{\sum_{k=1}^N u_{ik}^m}$$

With :

* $x_k$ the values of the observation $k$
* $v_i$ the values of the center of the cluster $i$
* $c$ the number of clusters
* $m$ the fuzzyness index

```{r warning=FALSE}
Cmean <- CMeans(Data,4,1.5,500,standardize = FALSE, seed = 456, tol = 0.00001, verbose = FALSE)
```

We can now use the function *calcqualityIndexes* which combines many indices from the package **fclust** to analyze the quality of the classification. We will use these values later to do comparisons among the different algorithms.

```{r message=FALSE, warning=FALSE}
calcqualityIndexes(Data, Cmean$Belongings, m = 1.5)
```

Note, **geocmeans** proposes also a generalized version of the c-means algorithm. It is known to accelerate convergence and yield less fuzzy results by adjusting the membership matrix at each iteration. It requires an extra parameter *beta* controlling the strength of the modification. The modification affects only the formula updating the membership matrix.

$$u_{ik} = \frac{(||x_{k} - v{_i}||^{2} - \beta_k) ^{(-1/(m-1))}}{\sum_{j=1}^c(||x_{k} - v{_j}||^2 - \beta_k)^{(-1/(m-1))}}$$

with $\beta_k = min(||x_{k} - v||^2)$ and $0 \leq \beta \leq 1$.

To select an appropriate value for this parameter, we will try all the possible values between 0 and 1 with a step of 0.05.

```{r warning=FALSE, eval = FALSE}
beta_values <- select_parameters("GFCM",data = Data, k = 4, m = 1.5,
                                 beta = seq(0,1,0.05), spconsist = F,
                                 tol = 0.00001, seed = 456)
```

```{r warning=FALSE}
knitr::kable(beta_values[c("beta","Silhouette.index","XieBeni.index","Explained.inertia")],
             col.names = c("beta", "silhouette index",
                           "Xie and Beni index", "explained inertia"),digits = 3)
```

Considering the table above, we select beta = 0.7, it keeps a good silhouhette index, and increases Xie and Beni index and explained inertia. Let us compare the results of GFCM and FCM.

```{r warning=FALSE}
GCmean <- GCMeans(Data,k = 4,m = 1.5, beta = 0.7,500,standardize = FALSE, seed=456,
                  tol = 0.00001, verbose = FALSE)
r1 <- calcqualityIndexes(Data,GCmean$Belongings,m=1.5)
r2 <- calcqualityIndexes(Data,Cmean$Belongings,m=1.5)
df <- cbind(unlist(r1), unlist(r2))

knitr::kable(df,
             digits = 3,col.names = c("GFCM", "FCM"))
```
The results indicate that the GFCM gives a solution that is less fuzzy (higher explained inertia and lower partition entropy) but keeps a good silhouette index and an even better Xie and Beni index.

We can now map the two membership matrices and the most likely group for each observation. To do so, we use the function *mapClusters* from **geocmeans**. We propose here to define a threshold of 0.45. If an observation has only values below this probability in a membership matrix, it will be labeled as "undecided" (represented with transparency on the map).

We can compare the maps of the classical c-means and the generalized version.
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 1", out.width="80%", fig.pos="H", fig.align="center"}
cmeansMaps<- mapClusters(LyonIris,Cmean$Belongings,undecided = 0.45)
GcmeansMaps<- mapClusters(LyonIris,GCmean$Belongings,undecided = 0.45)

ggarrange(cmeansMaps$ProbaMaps[[1]],GcmeansMaps$ProbaMaps[[1]], 
          nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 2", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(cmeansMaps$ProbaMaps[[2]],GcmeansMaps$ProbaMaps[[2]], 
          nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 3", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(cmeansMaps$ProbaMaps[[3]],GcmeansMaps$ProbaMaps[[3]], 
          nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 4", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(cmeansMaps$ProbaMaps[[4]],GcmeansMaps$ProbaMaps[[4]], 
          nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Most likely clusters and undecided units", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(cmeansMaps$ClusterPlot,GcmeansMaps$ClusterPlot,
          nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")
```

As expected, the results are very similar, but the generalized version provides a more clear-cut classification.

## Spatial c-means and generalized c-means

### Selecting *alpha* for SFCM

Now we can use the *SFCM* function to perform a spatial c-means. The first step is to define a spatial weight matrix indicating which observations are neighbours and the strength of their relationship. We propose here to use a basic queen neighbour matrix (built with **spdep**). The matrix must be row-standardized to ensure that the interpretation of all the parameters remains clear.

```{r message=FALSE, warning=FALSE}
library(spdep)

Neighbours <- poly2nb(LyonIris,queen = TRUE)
WMat <- nb2listw(Neighbours,style="W",zero.policy = TRUE)
```

The main challenge with the SFCM method is to select the parameter *alpha*. It represents the weight of the spatial dimension (lagged values) in the calculus of the membership matrix and the cluster centers.

* If alpha=0, then we end up with a classical c-means algorithm.
* If alpha=1, then the original and the lagged values have the same weight
* If alpha=2, then the lagged values are twice more important than the original values
* end so on...

The two following formulas show how the functions updating the membership matrix and the centers of the clusters are modified.

$$u_{ik} = \frac{(||x_{k} - v{_i}||^2 + \alpha||\bar{x_{k}} - v{_i}||^2)^{(-1/(m-1))}}{\sum_{j=1}^c(||x_{k} - v{_j}||^2 + \alpha||\bar{x_{k}} - v{_j}||^2)^{(-1/(m-1))}}$$

$$v_{i} = \frac{\sum_{k=1}^N u_{ik}^m(x_{k} + \alpha\bar{x_{k}})}{(1 + \alpha)\sum_{k=1}^N u_{ik}^m}$$
with $\bar{x}$ the lagged version of x and $\alpha \geq 0$

As the formula suggests, the *SFCM* can be see as a spatially smoothed version of the classical c-means and *alpha* controls the degree of spatial smoothness. This smoothing can be interpreted as an attempt to reduce spatial overfitting of the classical c-means.

To select *alpha*, we propose to check all the possible values between 0 and 2 with a step of 0.05.

```{r warning=FALSE,  message=FALSE, eval = FALSE}
DFindices_SFCM <- select_parameters(algo = "SFCM", data = Data,
                               k = 4, m = 1.5, alpha = seq(0,2,0.05),
                               nblistw = WMat, standardize = FALSE,
                               tol = 0.0001, verbose = FALSE, seed = 456)

```

Now we can check the indices to select the best *alpha*. The goal is to reduce as much as possible spatial inconsistency and to keep a good classification quality.

Let's start with the spatial inconsistency. This indicator (developed for this package) calculates the sum of the squared differences between each observation and its neighbours on the membership matrix. Thus, the maximum for each observation is $k*j$ with *j* the number of neighbours for the observation and *k* the number of groups. A maximum is reached if each observations have 100% chance to belong to a cluster that is different from all its neighbours. So, when we sum up the values obtained for all the observations, we obtain a quantity of spatial inconsistency. This quantity is divided by the quantity obtained when permuting randomly the rows of the membership matrix. This second quantity represents the spatial inconsistency that we might expect if the observations were randomly scattered in space. We can repeat the permutation step (Monte Carlo approach) and keep the mean of the ratios to have a more robust indicators (see help(spConsistency) for details).

A smaller value indicates a smaller spatial inconsistency and thus a greater spatial consistency. 0 meaning that all observations have exactly the same values in the membership matrix than their neighbours (perfect spatial consistency).

```{r warning=FALSE, fig.cap = "Link between alpha and spatial inconsistency", out.width="50%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFCM)+
  geom_smooth(aes(x=alpha,y=spConsistency), color = "black")+
  geom_point(aes(x=alpha,y=spConsistency), color = "red")

```

Unsurprisingly, increasing *alpha* leads to a decrease of the spatial inconsistency. This gain follow an inverse function.

Let's check now the explained inertia
```{r warning=FALSE, fig.cap = "Link between alpha and explained inertia", out.width="50%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFCM)+
  geom_smooth(aes(x=alpha,y=Explained.inertia), color = "black")+
  geom_point(aes(x=alpha,y=Explained.inertia), color = "red")
```

As expected, the explained inertia decreases when alpha increases and follows again an inverse function. The classification has to find a compromise between the original values and the lagged values. However, the loss is very small here: only 3% between alpha = 0 and alpha = 2.

To finish here, we can observe the silhouette and Xie and Beni indicators.

```{r fig.cap="Link between alpha and silhouette index", message=FALSE, warning=FALSE, fig.pos="H", fig.align="center", out.width="50%"}
ggplot(DFindices_SFCM)+
  geom_smooth(aes(x=alpha,y=Silhouette.index), color = "black")+
  geom_point(aes(x=alpha,y=Silhouette.index), color = "red")
```

```{r fig.cap="Link between alpha and Xie and Beni index", message=FALSE, warning=FALSE, out.width="50%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFCM)+
  geom_smooth(aes(x=alpha,y=XieBeni.index), color = "black")+
  geom_point(aes(x=alpha,y=XieBeni.index), color = "red")

```

The detail of the meaning of theses indicators is beyond the scope of this vignette. Let's just stress that a bigger silhouette index indicates a better classification, and a smaller Xie and beni index indicates a better classification.

After considering all the previous charts, we decide to keep alpha = 0.7 as it seems to give a good balance between spatial consistency and classification quality in this case.

```{r warning=FALSE}
SFCM <- SFCMeans(Data, WMat, k = 4, m = 1.5, alpha = 0.7,
                 tol = 0.0001, standardize = FALSE,
                 verbose = FALSE, seed = 456)
```

### Selecting alpha and beta for SGFCM

It is also possible to use the generalized version of the spatial c-means. In that case, we have to define both *alpha* and *beta*.

The next formula shows how the membership matrix is updated at each iteration. Note that the centers of the clusters are updated with the same formula as SFCM.

$$u_{ik} = \frac{(||x_{k} - v{_i}||^2 -\beta_k + \alpha||\bar{x_{k}} - v{_i}||^2)^{(-1/(m-1))}}{\sum_{j=1}^c(||x_{k} - v{_j}||^2 -\beta_k + \alpha||\bar{x_{k}} - v{_j}||^2)^{(-1/(m-1))}}$$

Because we select a high resolution for our grid search of *alpha* and *beta*, we will use a multiprocessing approach.

```{r warning=FALSE, eval = FALSE}
future::plan(future::multiprocess(workers=4))
DFindices_SFGCM <- select_parameters.mc(algo = "SGFCM", data = Data,
                               k = 4, m = 1.5, alpha = seq(0,2,0.05),
                               beta = seq(0,0.85,0.05),
                               nblistw = WMat, standardize = FALSE, chunk_size = 50,
                               tol = 0.0001, verbose = FALSE, seed = 456)
```

```{r warning=FALSE, fig.cap = "Impact of beta and alpha on silhouette index", out.width = "80%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFGCM) + 
  geom_raster(aes(x = alpha, y = beta, fill = Silhouette.index),  size = 5) + 
  scale_fill_viridis() +
  coord_fixed(ratio=1)
```
```{r warning=FALSE, fig.cap = "Impact of beta and alpha on Xie and Beni index", out.width = "80%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFGCM) + 
  geom_raster(aes(x = alpha, y = beta, fill = XieBeni.index),  size = 5) + 
  scale_fill_viridis() +
  coord_fixed(ratio=1)
```
```{r warning=FALSE, fig.cap = "Impact of beta and alpha on spatial inconsistency", out.width = "80%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFGCM) + 
  geom_raster(aes(x = alpha, y = beta, fill = spConsistency),  size = 5) + 
  scale_fill_viridis() +
  coord_fixed(ratio=1)

```

The first two plots indicate that some particular pairs of *alpha* and *beta* yield good results in the range 0.8 < *alpha* < 1.2 and 0.6 < *beta* < 0.8. The last plot shows that the selection of *beta* has no impact on the spatial consistency.

Considering the previous plots, we decide to retain the solution with *beta* = 0.65 and *alpha* = 0.95 which yield very good results for all the considered indices.

```{r warning=FALSE}
SGFCM <- SGFCMeans(Data,WMat,k = 4,m=1.5, alpha=0.95, beta = 0.65,
                    tol=0.0001, standardize = FALSE, verbose = FALSE, seed = 456)
```

Again, we compare here the generalized and the classical version of the spatial c-means algorithm.

```{r warning=FALSE}
r1 <- calcqualityIndexes(Data, SFCM$Belongings,m = 1.5)
r2 <- calcqualityIndexes(Data, SGFCM$Belongings,m = 1.5)

diagSFCM <- spatialDiag(belongmatrix = SFCM$Belongings, nblistw = WMat,
                        undecided = 0.45,nrep = 500)
diagSGFCM <- spatialDiag(belongmatrix = SGFCM$Belongings, nblistw = WMat,
                         undecided = 0.45,nrep = 500)

df <- cbind(
  c(unlist(r1),diagSFCM$SpConsist),
  c(unlist(r2),diagSGFCM$SpConsist)
)
row.names(df)[length(row.names(df))] <- "sp.consistency"

knitr::kable(df,digits = 3,col.names = c("SFCM","SGFCM"))
```

The solution of the SGFCM is better on the semantic and the spatial aspects.

We can compare the maps
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 1", out.width="80%", fig.pos="H", fig.align="center"}
SFCMMaps <- mapClusters(geodata = LyonIris,belongmatrix = SFCM$Belongings,undecided = 0.45)
SGFCMMaps <- mapClusters(geodata = LyonIris,belongmatrix = SGFCM$Belongings,undecided = 0.45)

ggarrange(SFCMMaps$ProbaMaps[[1]],SGFCMMaps$ProbaMaps[[1]], nrow = 1, ncol = 2,
          common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 2", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(SFCMMaps$ProbaMaps[[2]],SGFCMMaps$ProbaMaps[[2]], nrow = 1, ncol = 2,
          common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 3", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(SFCMMaps$ProbaMaps[[3]],SGFCMMaps$ProbaMaps[[3]], nrow = 1, ncol = 2,
          common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 4", out.width="80%"}
ggarrange(SFCMMaps$ProbaMaps[[4]],SGFCMMaps$ProbaMaps[[4]], nrow = 1, ncol = 2,
          common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Most likely cluster and undecided units", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(SFCMMaps$ClusterPlot,SGFCMMaps$ClusterPlot, nrow = 1, ncol = 2,
          common.legend = TRUE, legend = "bottom")
```

## Comparing the spatial consistency of FCM, GFCM, SFCM and SGFCM

Now, we can  do a deeper spatial analysis and compare the spatial consistency of the four classifications realized (FCM, GFCM, SFCM, SGFCM).

```{r warning=FALSE}
spdiag_1 <- spatialDiag(Cmean$Belongings, nblistw = WMat,nrep=250)
spdiag_2 <- spatialDiag(GCmean$Belongings, nblistw = WMat,nrep=250)
spdiag_3 <- spatialDiag(SFCM$Belongings, nblistw = WMat,nrep=250)
spdiag_4 <- spatialDiag(SGFCM$Belongings, nblistw = WMat,nrep=250)

#looking at the moran I values for each group
moran_table <- data.frame(cbind(spdiag_1$MoranValues$MoranI,
                     spdiag_2$MoranValues$MoranI,
                     spdiag_3$MoranValues$MoranI,
                     spdiag_4$MoranValues$MoranI
                     ))
row.names(moran_table) <- paste("cluster ",1:4,sep="")
knitr::kable(moran_table, digits = 3,
             col.names = c("FCM","GFCM","SFCM","SGFCM"),
             caption = "Moran I index for the columns of the membership matrix"
             )
```

Unsurprisingly, the Moran I values calculated on the membership matrices are higher for SFCM and SGFCM, indicating stronger spatial structures in the classifications.

```{r warning=FALSE}
print(c(spdiag_1$SpConsist, spdiag_2$SpConsist,spdiag_3$SpConsist,spdiag_4$SpConsist))
```

Considering the values of spatial inconsistency, we could check if the value obtained for SGFCM is significantly lower than the one of SFCM. Considering the previous `r length(spdiag_3$SpConsistSamples)` permutated values, we can calculate a pseudo p-value:

```{r warning=FALSE}
sum(spdiag_4$SpConsist > spdiag_3$SpConsistSamples) / length(spdiag_3$SpConsistSamples)
```

It appears that out of `r length(spdiag_3$SpConsistSamples)` permutations, the observed values of spatial inconsistency of SGFCM is always lower than the one of SFCM. The difference is significant at the threshold `r 1/length(spdiag_3$SpConsistSamples)`=1/`r length(spdiag_3$SpConsistSamples)`

We can map the undecided observations of the final solution. These entities should be analyzed more precisely. Selecting them is easy with the function *undecidedUnits* of the **geocmeans** package.

```{r warning=FALSE}
Undecided <- undecidedUnits(SGFCM$Belongings,0.45)
LyonIris$FinalCluster <- ifelse(Undecided=="Undecided",
                                "Undecided",paste("cluster",Undecided,sep="_"))

#mapping the groups
DFmapping <- merge(FortiData,LyonIris,by.x="id",by.y="OID")

ggplot(data=DFmapping)+
  geom_polygon(aes(x=long,y=lat,group=group,fill=FinalCluster),color=rgb(0,0,0,0))+
  coord_fixed(ratio = 1)+
  scale_fill_manual(name="FinalCluster",values = c("cluster_V1"="palegreen3",
                                 "cluster_V2"="firebrick",
                                 "cluster_V3"="lightyellow2",
                                 "cluster_V4"="steelblue",
                                 "cluster_V5"="pink",
                                 "Undecided"=rgb(0,0,0,0.4)))+
  theme( axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
    )
```

## Interpreting the results of the final classification

One can obtain many descriptive information about the final groups with three functions of **geocmeans** :

* *summarizeClusters* : calculate summary statistics for each group for a given dataset by using the membership matrix as weights (sticking with the fuzzy spirit).
* *spiderPlots* : display a spider plot allowing to compare quickly the differences between groups.
* *violinPlots* : display a violin plot for each variable in a given dataset. Observations must be grouped before.

```{r warning=FALSE}
summarizeClusters(LyonIris@data[AnalysisFields],belongmatrix = SGFCM$Belongings,
                  weighted = T,dec = 3)
```

```{r warning=FALSE, eval = FALSE}
spiderPlots(LyonIris@data[AnalysisFields], SGFCM$Belongings,
            chartcolors = c("darkorange3","grey4","darkgreen","royalblue"))
violinPlots(LyonIris@data[AnalysisFields], SGFCM$Groups)
```

That's all folks ! These are the enhancements for the next version

* introduce other methods of spatial c-means
* open some other parameters to the user (like the function defining the convergence criterion)
* work on documentation
* improve calculus speed by dropping some apply in the code
* adding the manhattan distance (useful when data have a high dimensionality)

### 2021-04-13 doc/introduction.R

@@ -1,0 +1,286 @@
+## ----setup, include=FALSE-----------------------------------------------------
+knitr::opts_chunk$set(echo = TRUE)
+
+## ----message=FALSE, warning=FALSE---------------------------------------------
+#charging packages and data
+library(geocmeans)
+library(ggplot2)
+library(ggpubr)
+library(dplyr)
+library(viridis)
+library(spdep)
+
+data(LyonIris)
+
+#selecting the columns for the analysis
+AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14",
+                   "Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")
+
+#rescaling the columns
+Data <- LyonIris@data[AnalysisFields]
+for (Col in names(Data)){
+  Data[[Col]] <- scale(Data[[Col]])
+}
+
+#preparing some elements for further mapping
+LyonIris$OID <- as.character(1:nrow(LyonIris))
+FortiData <- broom::tidy(LyonIris,region="OID")
+
+## ----include=FALSE------------------------------------------------------------
+#loading the pre-calculated results
+load(system.file("extdata", "results_vignette_intro.rda",
+                           package = "geocmeans", mustWork = TRUE))
+
+## ----warning=FALSE, fig.cap = "Impact of the number of groups on the explained variance", out.width = "50%", fig.pos="H", fig.align="center"----
+#finding the best k by using the r2 of the classification
+#trying for k from 2 to 10
+R2s <- sapply(2:10,function(k){
+  Clust <- kmeans(Data,centers=k,iter.max = 150)
+  R2 <- Clust$betweenss / Clust$totss
+  return(R2)
+})
+
+Df <- data.frame(K=2:10,
+                 R2 = R2s)
+
+ggplot(Df)+
+  geom_line(aes(x=K,y=R2s))+
+  geom_point(aes(x=K,y=R2s),color="red")+
+  xlab("Number of groups")+
+  ylab("R2 of classification")
+
+## ----warning=FALSE, fig.cap="Clusters identified by classical k-means", fig.pos="H", fig.align="center"----
+KMeanClust <-  kmeans(Data,centers=4,iter.max = 150)
+LyonIris$Cluster <-paste("cluster",KMeanClust$cluster,sep="_")
+
+#mapping the groups
[…]

### 2021-04-14 doc/introduction.Rmd
corrections in the vignettes

@@
 
-@cai2007fast described in their article the method, originally applied in the analysis of brain imagery. The generalized version of the spatial fuzzy c-means is presented by @zhao2013kernel.
+In their article, @cai2007fast described the method, originally applied in the analysis of brain imagery. The generalized version of the spatial fuzzy c-means is presented by @zhao2013kernel.
 
-@gelb2021apport applied the method to socio-residential and environmental data and compared the results with oter unsupervised classification algorithms (in french).
+@gelb2021apport applied the method to socio-residential and environmental data comparing the results with other unsupervised classification algorithms (in French).
 
 There are actually numerous packages and functions to perform unsupervised classification in R (*hclust*, *kmeans*, *cmeans*, *factoextra*, etc.).
-However theses methods are not always well-suited to analyze spatial data.
-Indeed, they do not take into account spatial information like proximity or contiguity between observations.
-This may lead to solutions for which close observations end up in different groups despite the fact that they are very similar.
+However, these methods are not always well-suited to analyze spatial data.
+Indeed, they do not account for spatial information such as proximity or contiguity between observations.
+This may lead to solutions for which close observations end up in different groups event though they are very similar.
 
-The package **ClustGeo** is to our knowledge the only package proposing an unsupervised classification method considering directly spatial proximity between observations.
-The proposed approach is very appealing because the user has to select a parameter (*alpha*) that controls the weight of the spatial distance matrix (calculated between observations with their locations) VS the semantic distance matrix (calculated  between observations with their variables).
+To our knowledge, the package **ClustGeo** is the only package proposing an unsupervised classification method which directly consider spatial proximity between observations.
+The proposed approach is appealing because the user can select a parameter (*alpha*) that controls the weight of the spatial distance matrix (calculated between observations with their locations) versus the semantic distance matrix (calculated  between observations with their variables).
 
@@
 The main draw-back here is the difficulty to identify observations that are undecided, at the frontier of two clusters/groups.
-The soft or fuzzy clustering algorithms give more information because they calculate for each observation its "probability" of belonging to each group.
+The soft or fuzzy clustering algorithms provide more information because they calculate the "probability" of each observation to belong to each group.
 
@@
 
-Before starting the analysis, it is required to standardize the data because most of the calculus is based on Euclidean distance (we plan to also include Manhattan distance in a future release).
+Before starting the analysis, the data must be standardized because most of the calculus is based on Euclidean distance (we plan to also include Manhattan distance in a future release).
 
@@
 
-To explore the dataset and choose a number of cluster/groups (k) we propose to start with a classical kmeans.
+To explore the dataset and choose the right number of cluster/groups (k) we propose to start with a classical kmeans.
 
@@
 
-The classical c-means is a simple method to perform fuzzy unsupervised classification. The package **geocmeans** proposes the function *CMeans*. We set the fuzzyness degree (m) to 1.5.
+The classical c-means is a simple method to perform fuzzy unsupervised classification. The package **geocmeans** proposes the function *CMeans*. We set the fuzziness degree (m) to 1.5.
 
-We do not present here the algorithm, but only the two main formulas.
+We do not present the algorithm here, but only the two main formulas.
 
@@
 * $c$ the number of clusters
-* $m$ the fuzzyness index
+* $m$ the fuzziness index
 
@@
 We can now use the function *calcqualityIndexes* which combines many indices from the package **fclust** to analyze the quality of the classification.
-We will use these values later to do comparisons among the different algorithms.
+We will use these values later for the purpose of comparisons among the different algorithms.
 
@@
 
-Note, **geocmeans** proposes also a generalized version of the c-means algorithm.
+Note, **geocmeans** also proposes a so-called generalized version of the c-means algorithm.
 It is known to accelerate convergence and yield less fuzzy results by adjusting the membership matrix at each iteration.
-It requires an extra parameter *beta* controlling the strength of the modification.
-The modification affects only the formula updating the membership matrix.
+It requires an extra *beta* parameter controlling the strength of the modification.
+The modification only affects the formula updating the membership matrix.
 
@@
 
-Considering the table above, we select beta = 0.7, it keeps a good silhouhette index, and increases Xie and Beni index and explained inertia. Let us compare the results of GFCM and FCM.
+Considering the table above, we select *beta* = 0.7, it maintains a good silhouette index, and increases Xie and Beni index and explained inertia. Let us compare the results of GFCM and FCM.
 
@@
 ```
-The results indicate that the GFCM gives a solution that is less fuzzy (higher explained inertia and lower partition entropy) but keeps a good silhouette index and an even better Xie and Beni index.
+The results indicate that the GFCM provides a solution that is less fuzzy (higher explained inertia and lower partition entropy) but keeps a good silhouette index and an even better Xie and Beni index.
 
@@
 We propose here to define a threshold of 0.45.
-If an observation has only values below this probability in a membership matrix, it will be labeled as "undecided" (represented with transparency on the map).
+If an observation only has values below this probability in a membership matrix, it will be labeled as "undecided" (represented with transparency on the map).
 
@@
 
-As the formula suggests, the *SFCM* can be see as a spatially smoothed version of the classical c-means and *alpha* controls the degree of spatial smoothness.
+As the formula suggests, the *SFCM* can be seen as a spatially smoothed version of the classical c-means and *alpha* controls the degree of spatial smoothness.
 This smoothing can be interpreted as an attempt to reduce spatial overfitting of the classical c-means.
 
-To select *alpha*, we propose to check all the possible values between 0 and 2 with a step of 0.05.
+To select *alpha*, we propose to check all possible values between 0 and 2 with a step of 0.05.
 
@@
 
-Now we can check the indices to select the best *alpha*. The goal is to reduce as much as possible spatial inconsistency and to keep a good classification quality.
+Now we are able to check the indices to select the best *alpha*. The goal is to reduce spatial inconsistency as much as possible and to maintain a good classification quality.
 
-Let's start with the spatial inconsistency.
+Let us start with the spatial inconsistency.
 This indicator (developed for this package) calculates the sum of the squared differences between each observation and its neighbours on the membership matrix.
 Thus, the maximum for each observation is $k*j$ with *j* the number of neighbours for the observation and *k* the number of groups.
-A maximum is reached if each observations have 100% chance to belong to a cluster that is different from all its neighbours.
+A maximum is reached if each observation has 100% chance belonging to a cluster that is different from all its neighbours.
 So, when we sum up the values obtained for all the observations, we obtain a quantity of spatial inconsistency.
-This quantity is divided by the quantity obtained when permuting randomly the rows of the membership matrix.
+This quantity is divided by the quantity obtained when randomly permuting the rows of the membership matrix.
 This second quantity represents the spatial inconsistency that we might expect if the observations were randomly scattered in space.
-We can repeat the permutation step (Monte Carlo approach) and keep the mean of the ratios to have a more robust indicators (see help(spConsistency) for details).
+We can repeat the permutation step (Monte Carlo approach) and keep the mean of the ratios to have a more robust indicator (see help(spConsistency) for details).
 
 A smaller value indicates a smaller spatial inconsistency and thus a greater spatial consistency.
-0 meaning that all observations have exactly the same values in the membership matrix than their neighbours (perfect spatial consistency).
+0 meaning that all observations have exactly the same values in the membership matrix as their neighbours (perfect spatial consistency).
 
@@
 
-Unsurprisingly, increasing *alpha* leads to a decrease of the spatial inconsistency. This gain follow an inverse function.
+Not surprisingly, increasing *alpha* leads to a decrease of the spatial inconsistency. This gain follows an inverse function.
 
-Let's check now the explained inertia
+Let us now check the explained inertia
 ```{r warning=FALSE, fig.cap = "Link between alpha and explained inertia", out.width="50%", fig.pos="H", fig.align="center"}
@@
 
-As expected, the explained inertia decreases when alpha increases and follows again an inverse function.
+As expected, the explained inertia decreases when alpha increases and again follows an inverse function.
 The classification has to find a compromise between the original values and the lagged values.
@@
 
-The detail of the meaning of theses indicators is beyond the scope of this vignette.
-Let's just stress that a bigger silhouette index indicates a better classification, and a smaller Xie and beni index indicates a better classification.
+The detail of the meaning of these indicators is beyond the scope of this vignette.
+Let us just stress that a larger silhouette index indicates a better classification, and a smaller Xie and beni index indicates a better classification.
 
-After considering all the previous charts, we decide to keep alpha = 0.7 as it seems to give a good balance between spatial consistency and classification quality in this case.
+After considering all the previous charts, we decide to keep alpha = 0.7 as it seems to provide a good balance between spatial consistency and classification quality in this case.
 
@@
 
-It is also possible to use the generalized version of the spatial c-means. In that case, we have to define both *alpha* and *beta*.
+It is also possible to use the so-called generalized version of the spatial c-means. In that case, we must define both *alpha* and *beta*.
 
-The next formula shows how the membership matrix is updated at each iteration. Note that the centers of the clusters are updated with the same formula as SFCM.
+The next formula shows how the membership matrix is updated at each iteration. Note that the centres of the clusters are updated with the same formula as SFCM.
 
@@
 
-The first two plots indicate that some particular pairs of *alpha* and *beta* yield good results in the range 0.8 < *alpha* < 1.2 and 0.6 < *beta* < 0.8.
+The first two plots indicate that some pairs of *alpha* and *beta* in particulat yield good results in the range 0.8 < *alpha* < 1.2 and 0.6 < *beta* < 0.8.
 The last plot shows that the selection of *beta* has no impact on the spatial consistency.
 
-Considering the previous plots, we decide to retain the solution with *beta* = 0.65 and *alpha* = 0.95 which yield very good results for all the considered indices.
+Considering the previous plots, we decide to retain the solution with *beta* = 0.65 and *alpha* = 0.95 which yield very good results for all the indices considered.
 
@@
 
-Unsurprisingly, the Moran I values calculated on the membership matrices are higher for SFCM and SGFCM, indicating stronger spatial structures in the classifications.
+Not surprisingly, the Moran I values calculated on the membership matrices are higher for SFCM and SGFCM, indicating stronger spatial structures in the classifications.
 
@@
 
-It appears that out of `r length(spdiag_3$SpConsistSamples)` permutations, the observed values of spatial inconsistency of SGFCM is always lower than the one of SFCM.
-The difference is significant at the threshold `r 1/length(spdiag_3$SpConsistSamples)`=1/`r length(spdiag_3$SpConsistSamples)`
+It appears that out of `r length(spdiag_3$SpConsistSamples)` permutations, the observed values of spatial inconsistency of SGFCM are always lower than that of SFCM.
+The difference is significant at the threshold `r 1/length(spdiag_3$SpConsistSamples)` (=1/`r length(spdiag_3$SpConsistSamples)`)
 
@@
 
-One can obtain many descriptive information about the final groups with three functions of **geocmeans** :
+One can obtain a lot of descriptive information about the final groups with three functions of **geocmeans** :
 
@@
 
-That's all folks ! These are the enhancements for the next version
+That's all, folks ! Following are the enhancements for the next version
 
 * introduce other methods of spatial c-means
-* open some other parameters to the user (like the function defining the convergence criterion)
+* open some other parameters to the user (such as the function defining the convergence criterion)
 * work on documentation
-* improve calculus speed by dropping some apply in the code
+* improve calculus speed by dropping some *apply* in the code
 * adding the manhattan distance (useful when data have a high dimensionality)

### 2021-04-15 R/side_functions.R
Adding an alias for selectParameters, and correcting the vignette

@@ -925,9 +925,9 @@ eval_parameters <- function(algo, parameters, data, nblistw = NULL, standardize 
 }
 
 
-#' @title Select parameters for clustering algorithm
+#' @title Select parameters for a clustering algorithm
 #'
-#' @description Function to select the parameters of the classification
+#' @description Function to select the parameters for a clustering algorithm.
 #'
 #' @param algo A string indicating which method to use (FCM, GFCM, SFCM, SGFCM)
 #' @param data A dataframe with numeric columns
@@ -984,12 +984,29 @@ select_parameters <- function(algo,data,k,m,alpha = NA, beta = NA, nblistw=NULL,
 }
 
 
+#' @rdname select_parameters
+#' @examples
+#' \dontrun{
+#' data(LyonIris)
+#' AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img",
+#' "TxChom1564","Pct_brevet","NivVieMed")
+#' dataset <- LyonIris@data[AnalysisFields]
+#' queen <- spdep::poly2nb(LyonIris,queen=TRUE)
+#' Wqueen <- spdep::nb2listw(queen,style="W")
+#' #set spconsist to TRUE to calculate the spatial consistency indicator
+#' #FALSE here to reduce the time during package check
+#' values <- selectParameters(algo = "SFCM", dataset, k = 5, m = seq(2,3,0.1),
+#'     alpha = seq(0,2,0.1), nblistw = Wqueen, spconsist=FALSE)
+#' }
+#' @export
+selectParameters <- select_parameters
+
 
 #' @title Select parameters for clustering algorithm (multicore)
 #'
-#' @description Function to select the parameters of the classification alpha, k and m.
+#' @description Function to select the parameters for a clustering algorithm.
 #' This version of the function allows to use a plan defined with the package
-#' future to reduce calculation time
+#' future to reduce calculation time.
 #'
 #' @param algo A string indicating which method to use (FCM, GFCM, SFCM, SGFCM)
 #' @param data A dataframe with numeric columns
@@ -1082,6 +1099,28 @@ select_parameters.mc <- function(algo,data,k,m, alpha = NA, beta = NA, nblistw =
     dfIndices <- do.call(rbind,values)
[…]

### 2021-04-15 doc/introduction.R

@@ -76,7 +76,7 @@ Cmean <- CMeans(Data,4,1.5,500,standardize = FALSE, seed = 456, tol = 0.00001, v
 calcqualityIndexes(Data, Cmean$Belongings, m = 1.5)
 
 ## ----warning=FALSE, eval = FALSE----------------------------------------------
-#  beta_values <- select_parameters("GFCM",data = Data, k = 4, m = 1.5,
+#  beta_values <- selectParameters("GFCM",data = Data, k = 4, m = 1.5,
 #                                   beta = seq(0,1,0.05), spconsist = F,
 #                                   tol = 0.00001, seed = 456)
 
@@ -125,7 +125,7 @@ Neighbours <- poly2nb(LyonIris,queen = TRUE)
 WMat <- nb2listw(Neighbours,style="W",zero.policy = TRUE)
 
 ## ----warning=FALSE,  message=FALSE, eval = FALSE------------------------------
-#  DFindices_SFCM <- select_parameters(algo = "SFCM", data = Data,
+#  DFindices_SFCM <- selectParameters(algo = "SFCM", data = Data,
 #                                 k = 4, m = 1.5, alpha = seq(0,2,0.05),
 #                                 nblistw = WMat, standardize = FALSE,
 #                                 tol = 0.0001, verbose = FALSE, seed = 456)
@@ -160,7 +160,7 @@ SFCM <- SFCMeans(Data, WMat, k = 4, m = 1.5, alpha = 0.7,
 
 ## ----warning=FALSE, eval = FALSE----------------------------------------------
 #  future::plan(future::multiprocess(workers=4))
-#  DFindices_SFGCM <- select_parameters.mc(algo = "SGFCM", data = Data,
+#  DFindices_SFGCM <- selectParameters.mc(algo = "SGFCM", data = Data,
 #                                 k = 4, m = 1.5, alpha = seq(0,2,0.05),
 #                                 beta = seq(0,0.85,0.05),
 #                                 nblistw = WMat, standardize = FALSE, chunk_size = 50,

### 2021-04-17 README.rmd
Adding the code of conduct and guidelines for contributions. Also creating a rmd version of the READEME

---
output: rmarkdown::github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  fig.path = "man/figures/",
  out.width = "100%"
)
```

# geocmeans <img src='man/figures/geocmeans_logo.png' align="right" height="138.5" />
## An R package to perform Spatial Fuzzy C-means.

<!-- badges: start -->
    [![R-CMD-check](https://github.com/JeremyGelb/geocmeans/workflows/R-CMD-check/badge.svg)](https://github.com/JeremyGelb/geocmeans/actions)
<!-- badges: end -->

## Installation

You can install a development version of the `geocmeans` package using the command below.

```
remotes::install_github(repo = "JeremyGelb/geocmeans", build_vignettes = TRUE, force = TRUE)
```
## Authors

Jeremy Gelb, Laboratoire d'Équité Environnemental INRS (CANADA),  Email: jeremy.gelb@ucs.inrs.ca

Philippe Apparcio, Laboratoire d'Équité Environnemental INRS (CANADA), Email: philippe.apparicio@ucs.inrs.ca

## About the package

Provides functions to apply Spatial Fuzzy c-means 
Algorithm, visualize and interpret results. This method is 
well suited when the user wants to analyze data with a 
fuzzy clustering algorithm and to account for the spatial dimension of 
the dataset. Tests for estimating the spatial consistency and classification 
quality are proposed in addition. The algorithms were developped first for 
brain imagery as described in the articles of [Cai and al. 2007](https://doi.org/10.1016/j.patcog.2006.07.011) and [Zaho and al. 2013](https://doi.org/10.1016/j.dsp.2012.09.016). [Gelb and Apparicio](https://doi.org/10.4000/cybergeo.36414) proposed to apply the method to perform a socio-residential and environmental taxonomy in Lyon (France).

Approaches for visualising uncertainty in spatial data are presented in this package. These include the three approaches developed in [Lucchesi and Wikle (2017)](http://faculty.missouri.edu/~wiklec/LucchesiWikle2017Stat) and a fourth approach presented in [Kuhnert et al. (2018)](https://publications.csiro.au/publications/#publication/PIcsiro:EP168206).

#### Fuzzy classification algorithms

Four Fuzzy classification algorithms are proposed :

* FCM: Fuzzy C-Means, with the function `CMeans`
* GFCM: Generalized Fuzzy C-Means, with the function `GCMeans`
* SFCM: Spatial Fuzzy C-Means, with the function `SCMeans`
* SGFCM: Spatial Generalized Fuzzy C-Means, with the function `SGCMeans`

Each function return a membership matrix, the data used for the classification (scaled if required) and the centers of the clusters.

#### Parameter selections

The algorithms available require different parameters to be fixed by the user. The function `selectParameters` is a useful tool to compare the results of different combinations of parameters. A multicore version, `selectParameters.mc`, using a plan from the package `future` is also available to speed up the calculus.

#### Classification quality

Many indices of classification quality can be calculated with the function `calcqualityIndexes`:

* *Silhouette.index*: the silhouette index (fclust::SIL.F)
* *Partition.entropy*: the partition entropy index (fclust::PE)
* *Partition.coeff*: the partition entropy coefficient (fclust::PC)
* *Modified.partition.coeff*: the modified partition entropy coefficient (fclust::MPC)
* *XieBeni.index* : the Xie and Beni index (fclust::XB)
* *FukuyamaSugeno.index* : the Fukuyama and Sugeno index (geocmeans::calcFukuyamaSugeno)
* *Explained.inertia*: the percentage of total inertia explained by the solution

#### Interpretation

Several functions are also available to facilitate the interpration of the classification:

* summary statistics for each cluster: `summarizeClusters`
* spider charts: `spiderPlots`
* violin plots: `violinPlots`
* maps of the membership matrix: `mapClusters` (support polygon, points and polylines)

#### Spatial inconsistency

We proposed an index to quantify the spatial inconsistency of a classification ([Gelb and Apparicio](https://doi.org/10.4000/cybergeo.36414)). If in a classification close observations tend to belong to the same group, then the value of the index is close to 0. If the index is close to 1, then the belonging to groups is randomly distributed in space. A value higher than one can happen in the case of negative spatial autocorrelation. The index is described in the vignette `adjustinconsistency`. The function `spatialDiag` does a complete spatial diagnostic of the membership matrix resulting from a classification.

## Examples

Detailed examples are given in the vignette `introduction`

```
vignette("introduction","geocmeans")
```

## Testing

If you would like to install and run the unit tests interactively, include `INSTALL_opts = "--install-tests"` in the installation code.

```
remotes::install_github(repo = "JeremyGelb/geocmeans", build_vignettes = TRUE, force = TRUE, INSTALL_opts = "--install-tests")
testthat::test_package("geocmeans", reporter = "stop")
```

## Contribute

To contribute to `geocmeans`, please follow these [guidelines](CONTRIBUTING.md).

Please note that the `Vizumap` project is released with a [Contributor Code of Conduct](CONDUCT.md). By contributing to this project, you agree to abide by its terms.

## License

`Vizumap` version 1.2.0 is licensed under [GPLv3](LICENSE.md).

## History of Vizumap

Vizumap began as a visualisation project at the University of Missouri in 2016. Chris Wikle, professor of statistics, posed an interesting research question to Lydia Lucchesi, a student curious about data visualisation and R.

How do you include uncertainty on a map displaying areal data estimates?

Over the course of a year, they put together three methods for visualising uncertainty in spatial statistics: the bivariate choropleth map, the pixel map, and the glyph map. By mid-2017, there were maps, and there was a lot of R code, but there was not a tool that others could use to easily make these types of maps, too. That’s when statistician Petra Kuhnert recommended developing an R package. Over the course of a month, Petra and Lydia developed Vizumap (originally named VizU) at CSIRO Data61 in Canberra, Australia. Since then, the package has been expanded to include exceedance probability maps, an uncertainty visualisation method developed by Petra while working on a Great Barrier Reef (GBR) project.

Vizumap has been used to visualise the uncertainty of American Community Survey estimates, the prediction errors of sediment estimates in a GBR catchment, and most recently the [uncertainty of estimated locust densities in Australia](https://www.nature.com/articles/s41598-020-73897-1/figures/4). We would like to assemble a Vizumap gallery that showcases different applications of the package’s mapping methods. If you use Vizumap to visualise uncertainty, please feel free to send the map our way. We would like to see it!

## References

Kuhnert, P.M., Pagendam, D.E., Bartley, R., Gladish, D.W., Lewis, S.E. and Bainbridge, Z.T. (2018) [Making management decisions in face of uncertainty: a case study using the Burdekin catchment in the Great Barrier Reef, Marine and Freshwater Research](https://publications.csiro.au/publications/#publication/PIcsiro:EP168206), 69, 1187-1200, https://doi.org/10.1071/MF17237.

Lucchesi, L.R. and Wikle C.K. (2017) [Visualizing uncertainty in areal data with bivariate choropleth maps, map pixelation and glyph rotation](http://faculty.missouri.edu/~wiklec/LucchesiWikle2017Stat), Stat, https://doi.org/10.1002/sta4.150.

### 2021-04-17 tests/testthat/test_belongingmatrices.R

@@ -1,95 +1,67 @@
 context("belonging matrices")
 
-test_that("SFCM belonging matrices must have a rowSums equal to 1",{
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+#### Testing that the obtained membership matrices have a rowSums = 1
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+test_that("Membership matrices must have a rowSums equal to 1",{
   data(LyonIris)
   library(spdep)
   AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")
   dataset <- LyonIris@data[AnalysisFields]
   queen <- poly2nb(LyonIris,queen=TRUE)
   Wqueen <- nb2listw(queen,style="W")
-  result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
-  x <- round(1-rowSums(result$Belongings),6)
-  expect_equal(sum(x),0)
+  FCM <- CMeans(dataset, k = 5, m = 1.5, standardize = TRUE)
+  GFCM <- GCMeans(dataset, k = 5, m = 1.5, beta = 0.5, standardize = TRUE)
+  SFCM <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
+  SGFCM <- SGFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, beta = 0.5, standardize = TRUE)
+  x1 <- round(1-rowSums(FCM$Belongings),6)
+  x2 <- round(1-rowSums(GFCM$Belongings),6)
+  x3 <- round(1-rowSums(SFCM$Belongings),6)
+  x4 <- round(1-rowSums(SGFCM$Belongings),6)
+  expect_equal(sum(x1) + sum(x2) + sum(x3) + sum(x4),0)
 })
 
-test_that("SFCM belonging matrices must have a rowSums equal to 1, with high alpha",{
+
[…]

### 2021-04-17 README.rmd
Generated the readme.md from readme.rmd

@@
 * FCM: Fuzzy C-Means, with the function `CMeans`
-* GFCM: Generalized Fuzzy C-Means, with the function `GCMeans`
-* SFCM: Spatial Fuzzy C-Means, with the function `SCMeans`
-* SGFCM: Spatial Generalized Fuzzy C-Means, with the function `SGCMeans`
+* GFCM: Generalized Fuzzy C-Means, with the function `GFCMeans`
+* SFCM: Spatial Fuzzy C-Means, with the function `SFCMeans`
+* SGFCM: Spatial Generalized Fuzzy C-Means, with the function `SGFCMeans`
 

### 2021-04-17 README.rmd
testthat is added to suggested package (it was missing and not mentionned by rcmd check, weird)

@@
 
-Please note that the `Vizumap` project is released with a [Contributor Code of Conduct](CONDUCT.md). By contributing to this project, you agree to abide by its terms.
+Please note that the `geocmeans` project is released with a [Contributor Code of Conduct](CONDUCT.md). By contributing to this project, you agree to abide by its terms.
 
@@
 
-`Vizumap` version 1.2.0 is licensed under [GPLv3](LICENSE.md).
-
-## History of Vizumap
-
-Vizumap began as a visualisation project at the University of Missouri in 2016.
-Chris Wikle, professor of statistics, posed an interesting research question to Lydia Lucchesi, a student curious about data visualisation and R.
-
-How do you include uncertainty on a map displaying areal data estimates?
-
-Over the course of a year, they put together three methods for visualising uncertainty in spatial statistics: the bivariate choropleth map, the pixel map, and the glyph map.
-By mid-2017, there were maps, and there was a lot of R code, but there was not a tool that others could use to easily make these types of maps, too.
-That’s when statistician Petra Kuhnert recommended developing an R package.
-Over the course of a month, Petra and Lydia developed Vizumap (originally named VizU) at CSIRO Data61 in Canberra, Australia.
-Since then, the package has been expanded to include exceedance probability maps, an uncertainty visualisation method developed by Petra while working on a Great Barrier Reef (GBR) project.
-
-Vizumap has been used to visualise the uncertainty of American Community Survey estimates, the prediction errors of sediment estimates in a GBR catchment, and most recently the [uncertainty of estimated locust densities in Australia](https://www.nature.com/articles/s41598-020-73897-1/figures/4).
-We would like to assemble a Vizumap gallery that showcases different applications of the package’s mapping methods.
-If you use Vizumap to visualise uncertainty, please feel free to send the map our way.
-We would like to see it!
-
-## References
-
-Kuhnert, P.M., Pagendam, D.E., Bartley, R., Gladish, D.W., Lewis, S.E.
-and Bainbridge, Z.T.
-(2018) [Making management decisions in face of uncertainty: a case study using the Burdekin catchment in the Great Barrier Reef, Marine and Freshwater Research](https://publications.csiro.au/publications/#publication/PIcsiro:EP168206), 69, 1187-1200, https://doi.org/10.1071/MF17237.
-
-Lucchesi, L.R.
-and Wikle C.K.
-(2017) [Visualizing uncertainty in areal data with bivariate choropleth maps, map pixelation and glyph rotation](http://faculty.missouri.edu/~wiklec/LucchesiWikle2017Stat), Stat, https://doi.org/10.1002/sta4.150.
+`geocmeans` version 0.1.0 is licensed under [GPL2 License](LICENSE.txt).

### 2021-04-17 README.rmd
working on the readme, the badge rcmd-check was not present before

@@
 <!-- badges: start -->
-[![R-CMD-check](https://github.com/JeremyGelb/geocmeans/workflows/R-CMD-check/badge.svg)](https://github.com/JeremyGelb/geocmeans/actions)
+[![R-CMD-check](https://github.com/JeremyGelb/geocmeans/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JeremyGelb/geocmeans/actions/workflows/R-CMD-check.yaml)
 <!-- badges: end -->

### 2021-04-18 R/GFCM.R
Starting to work on the JOSS paper. Added a small feature : k++ selection method for the initial centers of clusters

@@ -56,6 +56,9 @@ calcFGCMBelongMatrix <- function(centers, data, m, beta ){
 #' @param standardize A boolean to specify if the variables must be centered and
 #'   reduced (default = True)
 #' @param verbose A boolean to specify if the messages should be displayed
+#' @param init A string indicating how the initial centers must be selected. "random"
+#' indicates that random observations are used as centers. "kpp" use a distance based method
+#' resulting in more dispersed centers at the beginning. Both of them are heuristic.
 #' @param seed An integer used for random number generation. It ensures that the
 #' start centers will be the same if the same integer is selected.
 #' @return A named list with :
@@ -72,7 +75,7 @@ calcFGCMBelongMatrix <- function(centers, data, m, beta ){
 #' "TxChom1564","Pct_brevet","NivVieMed")
 #' dataset <- LyonIris@data[AnalysisFields]
 #' result <- GCMeans(dataset,k = 5, m = 1.5, beta = 0.5, standardize = TRUE)
-GCMeans <- function(data, k, m, beta, maxiter = 500, tol = 0.01, standardize = TRUE, verbose = TRUE, seed = NULL) {
+GCMeans <- function(data, k, m, beta, maxiter = 500, tol = 0.01, standardize = TRUE, verbose = TRUE, init = "random", seed = NULL) {
   # standardize data if required
   if (standardize) {
     if (verbose){
@@ -88,7 +91,7 @@ GCMeans <- function(data, k, m, beta, maxiter = 500, tol = 0.01, standardize = T
                          k = k, m = m, beta = beta,
                          maxiter = maxiter, tol = tol,
                          standardize = standardize, verbose = verbose,
-                         seed = seed)
+                         seed = seed, init = init)
   return(results)
 }
 
@@ -189,6 +192,9 @@ calcSFGCMBelongMatrix <- function(centers, data, wdata, m, alpha, beta ){
 #' @param standardize A boolean to specify if the variable must be centered and
 #'   reduced (default = True)
 #' @param verbose A boolean to specify if the progress bar should be displayed
+#' @param init A string indicating how the initial centers must be selected. "random"
+#' indicates that random observations are used as centers. "kpp" use a distance based method
+#' resulting in more dispersed centers at the beginning. Both of them are heuristic.
[…]

### 2021-04-19 README.rmd
minor edit in readme
preparing for a submission to JOSS

@@
 
+## Contributors
+
 Philippe Apparcio, Laboratoire d'Équité Environnemental INRS (CANADA), Email: philippe.apparicio@ucs.inrs.ca
@@
 fuzzy clustering algorithm and to account for the spatial dimension of 
-the dataset. Tests for estimating the spatial consistency and classification 
-quality are proposed in addition. The algorithms were developped first for 
+the dataset. Indexes for measuring the spatial consistency and classification 
+quality are proposed in addition. The algorithms were developed first for 
 brain imagery as described in the articles of [Cai and al.

### 2021-04-20 README.rmd
minor corrections after first CRAN human check

@@
 
-To contribute to `geocmeans`, please follow these [guidelines](CONTRIBUTING.md).
+To contribute to `geocmeans`, please follow these [guidelines](https://github.com/JeremyGelb/geocmeans/blob/master/CONTRIBUTING.md).
 
-Please note that the `geocmeans` project is released with a [Contributor Code of Conduct](CONDUCT.md). By contributing to this project, you agree to abide by its terms.
+Please note that the `geocmeans` project is released with a [Contributor Code of Conduct](https://github.com/JeremyGelb/geocmeans/blob/master/CONDUCT.md).
+By contributing to this project, you agree to abide by its terms.
 
@@
 
-`geocmeans` version 0.1.0 is licensed under [GPL2 License](LICENSE.txt).
+`geocmeans` version 0.1.0 is licensed under [GPL2 License](https://github.com/JeremyGelb/geocmeans/blob/master/LICENSE.txt).

### 2021-04-21 README.rmd
released on CRAN !

@@
 [![R-CMD-check](https://github.com/JeremyGelb/geocmeans/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JeremyGelb/geocmeans/actions/workflows/R-CMD-check.yaml)
+```{r echo=FALSE, results='asis'}
+cat(
+  badge_github_version(color = "green"),
+  badge_cran_release(color = "blue"),
+  badge_cran_download(color = "blue", type = "grand-total")
+)
+```
 <!-- badges: end -->
@@
 ## Installation
+
+The stable version of `geocmeans` is available on CRAN. You can install it with the command below.
+```
+install.packages("geocmeans")
+```
 

### 2021-04-21 R/side_functions.R

@@ -359,7 +359,7 @@ mapClusters <- function(geodata, belongmatrix, undecided = NULL) {
 #' #No example provided, this is an internal function, use the general wrapper function mapClusters
 mapPolygons <- function(geodata, belongmatrix, undecided = NULL){
     belongmatrix <- as.data.frame(belongmatrix)
-    names(belongmatrix) <- gsub(" ", "", names(belongmatrix), fixed = T)
+    names(belongmatrix) <- gsub(" ", "", names(belongmatrix), fixed = TRUE)
     geodata@data <- belongmatrix
     geodata$OID <- as.character(1:nrow(geodata@data))
 
@@ -428,7 +428,7 @@ mapPolygons <- function(geodata, belongmatrix, undecided = NULL){
 #' #No example provided, this is an internal function, use the general wrapper function mapClusters
 mapLines <- function(geodata, belongmatrix, undecided = NULL){
     belongmatrix <- as.data.frame(belongmatrix)
-    names(belongmatrix) <- gsub(" ", "", names(belongmatrix), fixed = T)
+    names(belongmatrix) <- gsub(" ", "", names(belongmatrix), fixed = TRUE)
     geodata@data <- cbind(geodata@data, belongmatrix)
     invisible(capture.output(FortiData <- broom::tidy(geodata, region = "OID")))
 
@@ -497,7 +497,7 @@ mapLines <- function(geodata, belongmatrix, undecided = NULL){
 #' #No example provided, this is an internal function, use the general wrapper function mapClusters
 mapPoints <- function(geodata, belongmatrix, undecided = NULL){
     belongmatrix <- as.data.frame(belongmatrix)
-    names(belongmatrix) <- gsub(" ", "", names(belongmatrix), fixed = T)
+    names(belongmatrix) <- gsub(" ", "", names(belongmatrix), fixed = TRUE)
     geodata@data <- cbind(geodata@data, belongmatrix)
     Coords <- sp::coordinates(geodata)
     geodata$X__ <- Coords[,1]
@@ -575,7 +575,7 @@ mapPoints <- function(geodata, belongmatrix, undecided = NULL){
 #' Wqueen <- spdep::nb2listw(queen,style="W")
 #' result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
 #' summarizeClusters(dataset, result$Belongings)
[…]

### 2021-04-21 README.rmd
updating the README

@@
 )
+library(badger)
 ```

### 2021-05-25 README.rmd
removing a small error in readme

@@
 [Gelb and Apparicio](https://doi.org/10.4000/cybergeo.36414) proposed to apply the method to perform a socio-residential and environmental taxonomy in Lyon (France).
-
-Approaches for visualising uncertainty in spatial data are presented in this package.
-These include the three approaches developed in [Lucchesi and Wikle (2017)](http://faculty.missouri.edu/~wiklec/LucchesiWikle2017Stat) and a fourth approach presented in [Kuhnert et al.
-(2018)](https://publications.csiro.au/publications/#publication/PIcsiro:EP168206).
 

### 2021-06-05 README.rmd
adding a gif
big commit,
the main results are now S3 objects, adding generic functions predict and summary.

adding ellipsis to shiny app,

adding several new cluster quality indices

@@
 
-* *Silhouette.index*: the silhouette index (fclust::SIL.F)
-* *Partition.entropy*: the partition entropy index (fclust::PE)
-* *Partition.coeff*: the partition entropy coefficient (fclust::PC)
-* *Modified.partition.coeff*: the modified partition entropy coefficient (fclust::MPC)
-* *XieBeni.index* : the Xie and Beni index (fclust::XB)
-* *FukuyamaSugeno.index* : the Fukuyama and Sugeno index (geocmeans::calcFukuyamaSugeno)
+* *Silhouette.index*: the silhouette index (`fclust::SIL.F`)
+* *Partition.entropy*: the partition entropy index (`fclust::PE`)
+* *Partition.coeff*: the partition entropy coefficient (`fclust::PC`)
+* *Modified.partition.coeff*: the modified partition entropy coefficient (`fclust::MPC`)
+* *XieBeni.index*: the Xie and Beni index (`fclust::XB`)
+* *FukuyamaSugeno.index*: the Fukuyama and Sugeno index (`geocmeans::calcFukuyamaSugeno`)
+* *DavidBoulin.index*: the David-Boulin index (`geocmeans::calcDavidBouldin`)
+* *CalinskiHarabasz.index*: the Calinski-Harabasz index (`geocmeans::calcCalinskiHarabasz`)
+* *GD43.index* and *GD53.index*: two version of the generalized Dunn index (`geocmeans::calcGD43` and `calcGD45`)
+* *Negentropy.index*: the Negentropy Increment index (`geocmeans::calcNegentropyI`)
 * *Explained.inertia*: the percentage of total inertia explained by the solution
@@
 
-* summary statistics for each cluster: `summarizeClusters`
+* summary statistics for each cluster: `summarizeClusters` (also accessible with the generic function `summary`)
 * spider charts: `spiderPlots`
@@
 * maps of the membership matrix: `mapClusters` (support polygon, points and polylines)
+
+There is also a shiny app that can be used to go deeper in the result interpretation. It requires the packages `shiny`, `leaflet`, `bslib`, `plotly`, `shinyWidgets`, `car`.
+
+![Alt Text](https://raw.githubusercontent.com/JeremyGelb/geocmeans/master/.github/gif/app_viz.gif)
 

### 2021-06-05 R/GAclustering.R

@@ -1,0 +1,73 @@
+# # %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+# #### THIS SECTION IS UNDER DEVELOPMENT ####
+# # %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+#
+#
+# Gaclustering <- function(dataset, k, listw, m, inter.loss = 0.1, max.alpha = 4, nepochs = 100, nspawns = 100,
+#                          tol = 0.001, maxiter = 300, standardize = TRUE){
+#
+#   data(LyonIris)
+#   AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img",
+#   "TxChom1564","Pct_brevet","NivVieMed")
+#   dataset <- LyonIris@data[AnalysisFields]
+#   queen <- spdep::poly2nb(LyonIris,queen=TRUE)
+#   Wqueen <- spdep::nb2listw(queen,style="W")
+#
+#   listw <- Wqueen
+#   nspawns <- 100
+#   max.alpha <- 4
+#   tol = 0.001
+#   maxiter = 300
+#   standardize = TRUE
+#   k = 4
+#   m = 1.5
+#   ## step1 : calculating the spatial moran eigen values
+#   spVectors <- adespatial::mem(listw)
+#
+#   ## step 2 : define some prior parameters
+#   n <- ncol(spMat)
+#   nl <- 1:n
+#   nlprobs <- (1/log(nl+1)) / sum(1/log(nl+1))
+#
+#   l <- 1:n
+#   lprobs <- rep(1/n, times = n)
+#
+#   ## defining the original situation
+#   #select for each spawn the number of vectors to use
+#   sel_nl <- sample(nl, size = nspawns, prob = nlprobs, replace = TRUE)
+#   total_vecs <- sum(sel_nl)
+#
+#   #now select the ids of the vectors to use
+#   sel_vecs <- sample(l, size = total_vecs, prob = lprobs, replace = TRUE)
+#   sel_vecs <- split(sel_vecs,
+#                     f = rep(1:nspawns,times=sel_nl))
+#
+#   #and the values of alpha
+#   alphas <- seq(0.01,max.alpha, 0.01)
+#   alphas_probs <- rep(1/length(alphas), times = length(alphas))
[…]

### 2021-06-05 R/assembling.R

@@ -244,6 +244,9 @@ main_worker <- function(algo, ...){
   results <- c(list(Centers = centers, Belongings = newbelongmatrix,
                   Groups = Groups, Data = data),params)
 
+  ## setting the class of the results
+  results <- FCMres(results)
+
   return(results)
 
 }

### 2021-06-07 README.rmd
correcting a typo in README

@@
 * *CalinskiHarabasz.index*: the Calinski-Harabasz index (`geocmeans::calcCalinskiHarabasz`)
-* *GD43.index* and *GD53.index*: two version of the generalized Dunn index (`geocmeans::calcGD43` and `calcGD45`)
+* *GD43.index* and *GD53.index*: two version of the generalized Dunn index (`geocmeans::calcGD43` and `calcGD53`)
 * *Negentropy.index*: the Negentropy Increment index (`geocmeans::calcNegentropyI`)

### 2021-06-12 R/boostrap_clust_validation.R
minor corrections to the bootstrap function and in the vignette

@@ -24,9 +24,11 @@
 #' @param init A string indicating how the initial centers must be selected. "random"
 #' indicates that random observations are used as centers. "kpp" use a distance based method
 #' resulting in more dispersed centers at the beginning. Both of them are heuristic.
+#' @param verbose A boolean to specify if the progress bar should be displayed.
 #' @return  A list of two values: group_consistency, a dataframe indicating for each cluster its
 #' consistency accross simulations. group_centers a list with a dataframe for each cluster. The values
 #' in the dataframes are the centers of the clusters at each simulation.
+#' @export
 #' @examples
 #' \dontrun{
 #' data(LyonIris)
@@ -43,9 +45,9 @@
 #'
 #' Cmean <- CMeans(Data,4,1.5,500,standardize = FALSE, seed = 456, tol = 0.00001, verbose = FALSE)
 #'
-#' validation <- bstp_group_validation(Cmean, nsim = 1000, maxiter = 1000, tol = 0.01, init = "random")
+#' validation <- boot_group_validation(Cmean, nsim = 1000, maxiter = 1000, tol = 0.01, init = "random")
 #' }
-bstp_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.01, init = "random"){
+boot_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.01, init = "random", verbose = TRUE){
 
   ## calulating the lagged dataset if necessary -----------------------
   if(object$algo %in% c("SGFCM", "SFCM")){
@@ -57,6 +59,10 @@ bstp_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.0
   k <- ncol(object$Belongings)
 
   ## Starting the iteration of the boostraping -----------------------
+  if(verbose){
+    pb <- txtProgressBar(1, nsim, style = 3)
+    print("Starting the bootstrap iterations...")
+  }
 
   all_perm <- lapply(1:nsim, function(i){
 
@@ -81,6 +87,10 @@ bstp_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.0
       Belongings = results$Belongings,
       idx = idx
     )
+    if (verbose){
+      setTxtProgressBar(pb, i)
+    }
+
     return(su_res)
   })
 
@@ -89,6 +99,7 @@ bstp_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.0
[…]

### 2021-06-13 README.rmd
Adding a new feature to facilitate the comparison of two results by updating order of groups. Also corrected a ton of errors in texts

@@
 * *FukuyamaSugeno.index*: the Fukuyama and Sugeno index (`geocmeans::calcFukuyamaSugeno`)
-* *DavidBoulin.index*: the David-Boulin index (`geocmeans::calcDavidBouldin`)
+* *DavidBoudlin.index*: the David-Bouldin index (`geocmeans::calcDavidBouldin`)
 * *CalinskiHarabasz.index*: the Calinski-Harabasz index (`geocmeans::calcCalinskiHarabasz`)
@@
 
-Several functions are also available to facilitate the interpration of the classification:
+Several functions are also available to facilitate the interpretation of the classification:
 

### 2021-06-13 R/FCM.R

@@ -62,7 +62,7 @@ calcCentroids <- function(data, belongmatrix, m){
 
 #' @title C-means
 #'
-#' @description The clasical c-mean algorithm
+#' @description The classical c-mean algorithm
 #'
 #' @param data A dataframe with only numerical variable
 #' @param k An integer describing the number of cluster to find
@@ -240,7 +240,7 @@ calcSWFCCentroids <- function(data, wdata, belongmatrix, m, alpha) {
 #' @return A named list with
 #' \itemize{
 #'         \item Centers: a dataframe describing the final centers of the groups
-#'         \item Belongings: the final bmembership matrix
+#'         \item Belongings: the final membership matrix
 #'         \item Groups: a vector with the names of the most likely group for each observation
 #'         \item Data: the dataset used to perform the clustering (might be standardized)
 #' }

### 2021-06-13 R/assembling.R

@@ -90,7 +90,7 @@ evaluateMatrices <- function(mat1, mat2, tol) {
 
 #' @title kpp centers selection
 #'
-#' @description Select the inital centers of centroids by using the k++ approach
+#' @description Select the initial centers of centroids by using the k++ approach
 #' as suggested in this article: http://ilpubs.stanford.edu:8090/778/1/2006-13.pdf
 #'
 #' @param data The dataset used in the classification

### 2021-06-13 R/clustering_evaluation.R

@@ -7,7 +7,7 @@
 #' @description Calculate the Negentropy Increment index of clustering quality.
 #'
 #' @details
-#' The Negentropy Increment index \insertCite{da2020incremental}{geocmeans} is based on the assumption that a normally shapped cluster is more
+#' The Negentropy Increment index \insertCite{da2020incremental}{geocmeans} is based on the assumption that a normally shaped cluster is more
 #' desirable. It uses the difference between the average negentropy
 #' of all the clusters in the partition, and that of the  whole partition.
 #' a smaller value indicates a better partition. The formula is:
@@ -78,7 +78,7 @@ calcNegentropyI <- function(data, belongmatrix, centers){
 #' \deqn{\delta_{r}\left(\omega_{i}, \omega_{j}\right)=\frac{\sum_{l=1}^{n}\left\|\boldsymbol{x_{l}}-\boldsymbol{c_{i}}\right\|^{\frac{1}{2}} . u_{il}+\sum_{l=1}^{n}\left\|\boldsymbol{x_{l}}-\boldsymbol{c_{j}}\right\|^{\frac{1}{2}} . u_{jl}}{\sum{u_{i}} + \sum{u_{j}}}}
 #'
 #' where \emph{u} is the membership matrix and \eqn{u_{i}} is the column of
-#' \emph{u} desrbing the membership of the \emph{n} observations to cluster
+#' \emph{u} describing the membership of the \emph{n} observations to cluster
 #' \emph{i}. \eqn{c_{i}} is the center of the cluster \emph{i}.
 #'
 #' The denominator is a measure of the maximal dispersion of all clusters, given
@@ -328,7 +328,7 @@ calcCalinskiHarabasz <- function(data, belongmatrix, centers){
 #' @param data The original dataframe used for the clustering (n*p)
 #' @param belongmatrix A membership matrix (n*k)
 #' @param centers The centers of the clusters
-#' @param m The fuzzyness parameter
+#' @param m The fuzziness parameter
 #' @return A float : the Fukuyama and Sugeno index
 #' @export
 #' @examples
@@ -557,7 +557,7 @@ calcqualityIndexes <- function(data, belongmatrix, m, indices = c("Silhouette.in
 #'   the random distribution of the spatial inconsistency
 #' @return A named list with :
 #' \itemize{
-#'         \item MoranValues : the moran I values fo each column of the membership
+#'         \item MoranValues : the moran I values for each column of the membership
 #'          matrix (spdep::MoranI)
[…]

### 2021-06-13 R/uncertainty.R

@@ -27,7 +27,7 @@ undecidedUnits <- function(belongmatrix, tol = 0.1) {
 }
 
 
-#' @title Diveristy index
+#' @title Diversity index
 #'
 #' @description Calculate the diversity (or entropy) index.
 #'

### 2021-07-23 R/FCM.R
many changes and corrections for raster support and extended spatial analysis of results. Part1

@@ -69,7 +69,7 @@ CMeans <- function(data, k, m, maxiter = 500, tol = 0.01, standardize = TRUE, ve
 #'
 #' @details The implementation is based on the following article : \doi{10.1016/j.patcog.2006.07.011}.\cr
 #'
-#' the matrix of belonging (u) is calculated as follow \cr
+#' the membership matrix (u) is calculated as follow \cr
 #' \deqn{u_{ik} = \frac{(||x_{k} - v{_i}||^2 + \alpha||\bar{x_{k}} - v{_i}||^2)^{(-1/(m-1))}}{\sum_{j=1}^c(||x_{k} - v{_j}||^2 + \alpha||\bar{x_{k}} - v{_j}||^2)^{(-1/(m-1))}}}
 #'
 #' the centers of the groups are updated with the following formula

### 2021-07-23 R/FCMclass.R

@@ -5,6 +5,7 @@
 #' @param obj A list, typically obtained from functions CMeans, GCMeans, SFCMeans, SGFCMeans
 #' @keywords internal
 #' @return A object of class FCMres
+#' @export
 #' @examples
 #' #This is an internal function, no example provided
 FCMres <- function(obj){
@@ -27,6 +28,8 @@ FCMres <- function(obj){
                error = function(e)
                  print("Obj$Data must be coercible to a matrix with as.matrix"))
 
+  obj$k <- ncol(obj$Belongings)
+
 
   if("Groups" %in% attrs == FALSE){
     DF <- as.data.frame(obj$Belongings)
@@ -40,11 +43,13 @@ FCMres <- function(obj){
 
 #' @title is method for FCMres
 #'
-#' @description Check if an object can me considered as a FCMres object
+#' @description Check if an object can be considered as a FCMres object
 #'
 #' @param x A FCMres object, typically obtained from functions CMeans, GCMeans, SFCMeans, SGFCMeans
 #' @return A boolean, TRUE if x can be considered as a FCMres object, FALSE otherwise
 #'   group
+#' @importFrom methods is
+#' @method is FCMres
 #' @export
 #' @examples
 #' data(LyonIris)
@@ -54,7 +59,7 @@ FCMres <- function(obj){
 #' queen <- spdep::poly2nb(LyonIris,queen=TRUE)
 #' Wqueen <- spdep::nb2listw(queen,style="W")
 #' result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
-#' is.FCMres(result)
+#' is(result, "FCMres")
 is.FCMres <- function(x){
   attrs <- names(x)
 
@@ -84,5 +89,327 @@ is.FCMres <- function(x){
   # if we pass all these steps, then the object could be considered as
   # a FCMres
   return(TRUE)
+}
+
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+####                      SUMMARY S3 METHOD                          ####
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+
+#' @title Descriptive statistics by group
+#'
+#' @description Calculate some descriptive statistics of each group
+#'
[…]

### 2021-07-23 R/GFCM.R

@@ -71,7 +71,7 @@ GCMeans <- function(data, k, m, beta, maxiter = 500, tol = 0.01, standardize = T
 #'
 #' @details The implementation is based on the following article : \doi{10.1016/j.dsp.2012.09.016}.\cr
 #'
-#' the matrix of belonging (u) is calculated as follow \cr
+#' the membership matrix (u) is calculated as follow \cr
 #' \deqn{u_{ik} = \frac{(||x_{k} - v{_i}||^2 -b_k + \alpha||\bar{x_{k}} - v{_i}||^2)^{(-1/(m-1))}}{\sum_{j=1}^c(||x_{k} - v{_j}||^2 -b_k + \alpha||\bar{x_{k}} - v{_j}||^2)^{(-1/(m-1))}}}
 #'
 #' the centers of the groups are updated with the following formula

### 2021-07-23 R/RcppExports.R

@@ -70,7 +70,6 @@ calcSWFCCentroids <- function(data, wdata, belongmatrix, m, alpha) {
 #' @return A n * k matrix representing the belonging probabilities of each
 #'   observation to each cluster
 #' @keywords internal
-#' #This is an internal function, no example provided
 #'
 calcSFCMBelongMatrix <- function(centers, data, wdata, m, alpha) {
     .Call(`_geocmeans_calcSFCMBelongMatrix`, centers, data, wdata, m, alpha)
@@ -119,7 +118,7 @@ calcSFGCMBelongMatrix <- function(centers, data, wdata, m, alpha, beta) {
 #' @title focal euclidean distance on a matrix with a given window
 #' @name focal_euclidean_mat_window
 #' @param mat a matrix
-#' @param window a numeric matrix (squarred)
+#' @param window a numeric matrix (squared)
 #' @return a matrix with the euclidean distance of each cell to
 #' its neighbours.
 #' @keywords internal
@@ -140,6 +139,42 @@ focal_euclidean_mat_window <- function(mat, window) {
 #'
 focal_euclidean_list <- function(matrices, window) {
     .Call(`_geocmeans_focal_euclidean_list`, matrices, window)
+}
+
+#' @title focal euclidean distance on a matrix with a given window for a cube
+#' @name focal_euclidean_arr_window
+#' @param mat an array (cube)
+#' @param window a numeric matrix (squared)
+#' @return a matrix with the euclidean distance of each cell to
+#' its neighbours.
+#' @keywords internal
+#' @export
+#'
+focal_euclidean_arr_window <- function(mat, window) {
+    .Call(`_geocmeans_focal_euclidean_arr_window`, mat, window)
+}
+
+#' @title Jaccard similarity coefficient
+#' @name calc_jaccard_idx
+#' @description Calculate the Jaccard similarity coefficient
+#' @param x A vector of positive reals
+#' @param y A vector of positive reals
+#' @return A double: the Jaccard similarity coefficient
+#' @keywords internal
+calc_jaccard_idx <- function(x, y) {
+    .Call(`_geocmeans_calc_jaccard_idx`, x, y)
+}
+
+#' @title Jaccard similarity coefficient between columns of two matrices
+#' @name calc_jaccard_mat
+#' @description Calculate the Jaccard similarity coefficient between the
+#' columns of two matrices
+#' @param matX A matrix
[…]

### 2021-07-23 R/assembling.R

@@ -105,7 +105,7 @@ evaluateMatrices <- function(mat1, mat2, tol) {
 #' #This is an internal function, no example provided
 kppCenters <- function(data,k){
   # step1: select the first center at random
-  centers <- as.matrix(data[sample(1:nrow(data),size = 1),])
+  centers <- t(as.matrix(data[sample(1:nrow(data),size = 1),], nrow = 1))
   # select the other centers
   for (i in 2:k){
     dists <- apply(centers,1,function(ci){
@@ -325,33 +325,82 @@ sanity_check <- function(dots,data){
 # This set of function is used to provide the parameters in the right order
 # to the intermediate functions
 
+#' @title membership matrix calculator for FCM algorithm
+#' @param data a matrix (the dataset used for clustering)
+#' @param centers a matrix (the centers of the clusters)
+#' @param dots a list of other arguments specific to FCM
+#' @return a matrix with the new membership values
+#' @keywords internal
 belongsFCM <- function(data, centers ,dots){
   return(calcBelongMatrix(centers,data,dots$m))
 }
+#' @title center matrix calculator for FCM algorithm
+#' @param data a matrix (the dataset used for clustering)
+#' @param centers a matrix (the centers of the clusters)
+#' @param belongmatrix a matrix with the membership values
+#' @param dots a list of other arguments specific to FCM
+#' @return a matrix with the new centers
+#' @keywords internal
 centersFCM <- function(data, centers, belongmatrix, dots){
   return(calcCentroids(data,belongmatrix, dots$m))
 }
 
-
+#' @title membership matrix calculator for GFCM algorithm
+#' @param data a matrix (the dataset used for clustering)
+#' @param centers a matrix (the centers of the clusters)
+#' @param dots a list of other arguments specific to FCM
+#' @return a matrix with the new membership values
+#' @keywords internal
 belongsGFCM <- function(data,centers,dots){
   return(calcFGCMBelongMatrix(centers,data,dots$m,dots$beta))
 }
+#' @title center matrix calculator for GFCM algorithm
+#' @param data a matrix (the dataset used for clustering)
+#' @param centers a matrix (the centers of the clusters)
+#' @param belongmatrix a matrix with the membership values
+#' @param dots a list of other arguments specific to FCM
+#' @return a matrix with the new centers
+#' @keywords internal
 centersGFCM <- function(data, centers, belongmatrix, dots){
[…]

### 2021-07-23 R/boostrap_clust_validation.R

@@ -4,7 +4,7 @@
 #'
 #' @details Considering that the classification produced by a FCM like algorithm
 #'   depends on its initial state, it is important to check if the groups
-#'   obtained are stable. This function use a bootstrap method to do so. During
+#'   obtained are stable. This function uses a bootstrap method to do it. During
 #'   a selected number of iterations (at least 1000), a sample of size n (with
 #'   replacement) is drawn from the original dataset. For each sample, the same
 #'   classification algorithm is applied and the results are compared with the
@@ -97,17 +97,20 @@ boot_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.0
   }
 
   cons_values <- lapply(all_perm, function(results){
-    qual_mat <- matrix(-1, nrow = k, ncol = k)
-    for(ii in 1:k){
-      for(jj in 1:k){
-        if(qual_mat[ii,jj] == -1){
-          x <- results$Belongings[,ii]
-          y <- object$Belongings[,jj][results$idx]
-          val <- calc_jaccard_idx(x,y)
-          qual_mat[ii,jj] <- val
-        }
-      }
-    }
+    # qual_mat <- matrix(-1, nrow = k, ncol = k)
+    # for(ii in 1:k){
+    #   for(jj in 1:k){
+    #     if(qual_mat[ii,jj] == -1){
+    #       x <- results$Belongings[,ii]
+    #       y <- object$Belongings[,jj][results$idx]
+    #       val <- calc_jaccard_idx(x,y)
+    #       qual_mat[ii,jj] <- val
+    #     }
+    #   }
+    # }
+    matX <- results$Belongings
+    matY <- object$Belongings[results$idx,]
+    qual_mat <- calc_jaccard_mat(matX,matY)
     colnames(qual_mat) <- 1:ncol(qual_mat)
     rownames(qual_mat) <- 1:nrow(qual_mat)
 
@@ -214,7 +217,7 @@ boot_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.0
 #'
 #' future::plan(future::multiprocess(workers=2))
 #'
-#' validation.mc <- boot_group_validation(Cmean, nsim = 1000, maxiter = 1000,
+#' validation <- boot_group_validation.mc(Cmean, nsim = 1000, maxiter = 1000,
 #'     tol = 0.01, init = "random")
[…]

### 2021-07-23 R/moran_raster.R

@@ -1,66 +1,0 @@
-
-#' @title Global Moran I for raster
-#'
-#' @description Calculate the global Moran I for a numeric raster
-#'
-#' @param rast A RasterLayer or a matrix
-#' @param window The window defining the neighbour weights
-#' @return A float: the global Moran I
-#' @examples
-#' # this is an internal function, no example provided
-calc_moran_raster <- function(rast, window){
-  if(class(rast)[[1]] == "RasterLayer"){
-    mat <- raster::as.matrix(rast)
-  }else if (class(rast)[[1]] == "matrix"){
-    mat <- rast
-  }else{
-    stop("rast parameter must be on of matrix or RasterLayer")
-  }
-  if(class(window)[[1]] == "matrix"){
-    fun <- moranI_matrix_window
-  }else if (class(window)[[1]] == "numeric"){
-    fun <- moranI_matrix
-  }else{
-    stop("window parameter must be an integer or a matrix")
-  }
-  val <- fun(mat, window)
-  return(val)
-}
-
-
-#' @title Local Moran I for raster
-#'
-#' @description Calculate the Local Moran I for a numeric raster
-#'
-#' @param rast A RasterLayer or a matrix
-#' @param window The window defining the neighbour weights
-#' @return A vector: the local Moran I
-#' @examples
-#' # this is an internal function, no example provided
-calc_local_moran_raster <- function(rast, window){
-  if(class(rast)[[1]] == "RasterLayer"){
-    mat <- raster::as.matrix(rast)
-  }else if (class(rast)[[1]] == "matrix"){
-    mat <- rast
-  }else{
-    stop("rast parameter must be on of matrix or RasterLayer")
-  }
-  if(class(window)[[1]] == "matrix"){
-    window <- window
-  }else if (class(window)[[1]] == "numeric"){
-    w <- 1+2*window
-    window <- matrix(1, nrow = w, ncol = w)
-    window[ceiling(w/2),ceiling(w/2)] <- 0
-  }else{
-    stop("window parameter must be an integer or a matrix")
-  }
-  vals <- local_moranI_matrix_window(mat, window)
-
-  if(class(rast)[[1]] == "RasterLayer"){
-    raster::values(rast) <- vals
-    return(rast)
-  }else{
-    return(matrix(vals, ncol = ncol(mat), nrow = nrow(mat)))
-  }
-
-}

### 2021-07-23 R/raster_support.R

@@ -1,47 +1,3 @@
-# # https://rdrr.io/bioc/OLIN/man/ma.matrix.html
-# # https://www.rdocumentation.org/packages/raster/versions/3.4-10/topics/focal
-# # example de jeu de donnees : https://www.r-exercises.com/2018/02/28/advanced-techniques-with-raster-data-part-1-unsupervised-classification/?__cf_chl_jschl_tk__=703ae8c7214fe36addfe12006afebc7442de2077-1624418523-0-AQOrJJeKBwWikYZcyrzGFdsNhq2ERFu4EIyPoPuvPuVd77RyG-ppxAcA-KJizHAhNWTVXc7crst9Hs-IBtbOl1yfjSzVGcRFzlsof2sCbYRPk4F5GrKrhn64zm5fOGhiMi25RcoqrnuYbjLpu5QFzyVtBlWiSvQ_IF1SO27b388v_IXyW5G-BHA1htk5vJeHBHnUZDDmDQNUeVhRDzgVrZVjCSP1akW5P854dj0ILOKNcYoqrAQVxvpUaonarM38I6QiQaTcgTpnXxSPXkNYaOAmLkReEjcpYXTTT4gETRH5Ymnd8fDElBBq9pkmvKxjGc-Sx5-Y15o-iAGKenGYWQSdXrypzaML-mMAJSSM7GEgL2a6dJ1HRtOHcxwmVAW6MuUPR2cabYkEzyTlWgfZsWk_l92jaiaqo_7KCo92WZRxWk3s3f3CfKhLenCglEjXJ-F9Rkn-LDPD-Iuh4NDABe8YHHs_F7Gbrb9TL__ii50bNlx9rN03RBV4H4ZZcRaSiw
-
-# # # # # #TODO : add waiter package to have a nice waiting pannel !
-# library(raster)
-# raster1 <- raster(".Rproj.user/raster_data_test/estuary_LC.tif")
-#
-# data <- lapply(1:nbands(raster1), function(i){
-#   myband <- raster(".Rproj.user/raster_data_test/estuary_LC.tif", band = i)
-#   return(myband)
-# })
-#
-# #names(data) <- paste("band",1:nbands(raster1), sep="")
-#
-# window <- matrix(1, nrow = 3, ncol = 3)
-# window <- window / sum(window)
-#
-# lag_method  <- "sum"
-# standardize <- TRUE
-# na.rm <- TRUE
-#
-# k = 5
-# m = 1.5
[…]

### 2021-07-23 R/spatial_indices.R

@@ -1,0 +1,829 @@
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+#### spatial consistency index ####
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+#' @title Spatial consistency index
+#'
+#' @description Calculate a spatial consistency index
+#'
+#' @details This index is experimental, it aims to measure how much a clustering solution
+#' is spatially consistent. A classification is spatially inconsistent if
+#' neighbouring observation do not belong to the same group. See detail for
+#' a description of its calculation
+#'
+#' The total spatial inconsistency (*Scr*) is calculated as follow
+#'
+#' \deqn{isp = \sum_{i}\sum_{j}\sum_{k} (u_{ik} - u_{jk})^{2} * W_{ij}}
+#'
+#' With U the membership matrix, i an observation, k the neighbours of i and W
+#' the spatial weight matrix This represents the total spatial inconsistency of
+#' the solution (true inconsistency) We propose to compare this total with
+#' simulated values obtained by permutations (simulated inconsistency). The
+#' values obtained by permutation are an approximation of the spatial
+#' inconsistency obtained in a random context Ratios between the true
+#' inconsistency and simulated inconsistencies are calculated A value of 0
+#' depict a situation where all observations are identical to their neighbours
+#' A value of 1 depict a situation where all observations are as much different
+#' as their neighbours that what randomness can produce A classification
+#' solution able to reduce this index has a better spatial consistency
+#'
+#' @template FCMresobj-arg
+#' @template nblistw2-arg
+#' @param window if rasters were used for the classification, the window must be
+#' specified instead of a list.w object. Can also be NULL if object is a FCMres object.
+#' @param nrep An integer indicating the number of permutation to do to simulate
+#' spatial randomness. Note that if rasters are used, each permutation can be very long.
+#' @param adj A boolean indicating if the adjusted version of the indicator must be
+#' calculated when working with rasters. When working with vectors, see the function
+#' adjustSpatialWeights to modify the list.w object.
[…]

### 2021-07-23 man-roxygen/FCMresobj-arg.R
part2 of great cleaning

@@ -1,0 +1,2 @@
+#' @param object A FCMres object, typically obtained from functions CMeans,
+#'   GCMeans, SFCMeans, SGFCMeans. Can also be a simple membership matrix.

### 2021-07-23 man-roxygen/nblistw2-arg.R

@@ -1,0 +1,3 @@
+#' @param nblistw A list.w object describing the neighbours typically produced
+#'   by the spdep package. Required if data is a dataframe, see the parameter window
+#'   if you use a list of rasters as input. Can also be NULL if object is a FCMres object.

### 2021-07-23 src/focal_cpp.cpp

@@ -40,7 +40,7 @@ using namespace Rcpp;
 //' @title focal euclidean distance on a matrix with a given window
 //' @name focal_euclidean_mat_window
 //' @param mat a matrix
-//' @param window a numeric matrix (squarred)
+//' @param window a numeric matrix (squared)
 //' @return a matrix with the euclidean distance of each cell to
 //' its neighbours.
 //' @keywords internal
@@ -134,6 +134,89 @@ NumericMatrix focal_euclidean_list(List matrices, NumericMatrix window){
   return m2;
 }
 
+
+
+//' @title focal euclidean distance on a matrix with a given window for a cube
+//' @name focal_euclidean_arr_window
+//' @param mat an array (cube)
+//' @param window a numeric matrix (squared)
+//' @return a matrix with the euclidean distance of each cell to
+//' its neighbours.
+//' @keywords internal
+//' @export
+//'
+// [[Rcpp::export]]
+NumericMatrix focal_euclidean_arr_window(arma::cube mat, arma::mat window){
+
+  int nr = mat.n_rows;
+  int nc = mat.n_cols;
+  int nl = mat.n_slices;
+  int r,c,lr,hr,lc,hc,i,j;
+  int wr,wc;
+  int w1 = floor(float(window.n_rows)/2.0);
+  int w2 = floor(float(window.n_cols)/2.0);
+  int window_width = window.n_rows;
+  int window_height = window.n_cols;
+  int start_window_row = 0, end_window_row = window_width-1, start_window_col = 0, end_window_col = window_height-1;
+  arma::mat subwindow;
+  double distsums;
+  arma::vec xi,xj;
+  arma::cube subcube;
+  // creating an empty matrix
+  NumericMatrix m2(nr, nc);
+
+  for (r = 0; r < nr ; r++){
+    wr = w1;
+    lr = r - w1;
+    hr = r + w1;
+    start_window_row = 0;
+    end_window_row = window_width-1;
+    if(lr < 0){
+      wr = w1 + lr;
+      start_window_row = lr * -1;
+      lr = 0;
+    }
+    if(hr >= nr){
+      end_window_row = window_width - 2 - (hr - nr); // so the width of the window minus excess
+      hr = nr-1;
+    }
+    for(c = 0; c < nc; c++){
+      start_window_col = 0;
+      end_window_col = window_height-1;
+      wc = w2;
+      lc = c - w2;
+      hc = c + w2;
[…]

### 2021-07-23 src/jaccard_idx.cpp

@@ -1,0 +1,46 @@
+#include "shared_functions.h"
+#include <cmath>
+using namespace Rcpp;
+
+
+
+//' @title Jaccard similarity coefficient
+//' @name calc_jaccard_idx
+//' @description Calculate the Jaccard similarity coefficient
+//' @param x A vector of positive reals
+//' @param y A vector of positive reals
+//' @return A double: the Jaccard similarity coefficient
+//' @keywords internal
+// [[Rcpp::export]]
+double calc_jaccard_idx(arma::vec x, arma::vec y){
+  return arma::accu(arma::min(x,y))/arma::accu(arma::max(x,y));
+}
+
+
+//' @title Jaccard similarity coefficient between columns of two matrices
+//' @name calc_jaccard_mat
+//' @description Calculate the Jaccard similarity coefficient between the
+//' columns of two matrices
+//' @param matX A matrix
+//' @param matY A matrix
+//' @return A matrix with the Jaccard index values
+//' @keywords internal
+// [[Rcpp::export]]
+NumericMatrix calc_jaccard_mat(NumericMatrix matX, NumericMatrix matY){
+  int k = matY.ncol();
+  int i,j;
+  // creating an empty matrix at first
+  NumericMatrix jacc_mat(k,k);
+  jacc_mat.fill(-1);
+  for(i = 0; i < k; i++){
+    for(j = 0; j < k; j++){
+      if(jacc_mat(i,j)==-1){
+          jacc_mat(i,j) = calc_jaccard_idx(matX(_,j),matY(_,i));
+        };
+    }
+  }
+  return jacc_mat;
+}
+
+
+

### 2021-07-23 src/shared_functions.cpp

@@ -73,6 +73,21 @@ NumericVector calcEuclideanDistance2(NumericMatrix y, NumericVector x){
   return result;
 }
 
+//' @title euclidean distance between rows of a matrix and a vector (arma mode)
+//' @name calcEuclideanDistance3
+//' @param y a matrix
+//' @param x a vector (same length as ncol(matrix))
+//' @return a vector (same length as nrow(matrix))
+//' @export
+//' @keywords internal
+//'
+// [[Rcpp::export]]
+arma::mat calcEuclideanDistance3(arma::mat y, arma::mat x){
+  return arma::sum(arma::pow(y.each_row() - x,2),1);
+}
+
+
+
 //' @title sum of two matrices by column
 //' @name add_matrices_bycol
 //' @param x a matrix
@@ -248,4 +263,18 @@ LogicalMatrix test_inferior_mat(NumericMatrix mat, double t){
   return result;
 }
 
-
+//' @title create a matrix by multiplying a vector by its elements one by one as rows
+//' @name vector_out_prod
+//' @param x a vector
+//' @return a NumericMatrix
+//' @keywords internal
+//' @export
+//'
+// [[Rcpp::export]]
+NumericMatrix vector_out_prod(NumericVector x){
+  NumericMatrix res(x.length(),x.length());
+  for(int i = 0; i < x.length(); i++){
+      res(i,_) = x(i) * x;
+  }
+  return res;
+}

### 2021-07-23 src/shared_functions.h

@@ -1,12 +1,17 @@
+#include "geocmeans.h"
+
 #ifndef shared_functions
 #define shared_functions
 
-#include <Rcpp.h>
-using namespace Rcpp;
+// #include <Rcpp.h>
+// using namespace Rcpp;
+// using namespace arma;
 
 NumericMatrix power_mat(NumericMatrix x, double p);
 
 NumericVector calcEuclideanDistance2(NumericMatrix y, NumericVector x);
+
+arma::vec calcEuclideanDistance3(arma::mat y, arma::mat x);
 
 NumericMatrix add_matrices_bycol(NumericMatrix x, NumericMatrix y);
 
@@ -23,6 +28,8 @@ NumericMatrix pow_matrix_bycol(NumericMatrix x, float p);
 NumericVector rowmins_mat(NumericMatrix x);
 
 LogicalMatrix test_inferior_mat(NumericMatrix mat, double t);
+
+NumericMatrix vector_out_prod(NumericVector x);
 
 double max_mat(NumericMatrix x);
 

### 2021-07-23 tests/testthat/test_spatial_functions.R

@@ -1,0 +1,242 @@
+context("spatial function tests")
+library(raster)
+
+
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+#### Testing the global and local moran I
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+test_that("The global moran I calculated is the same as the one in raster package",{
+  mat <- matrix(sample(1:10, size = 100, replace = TRUE), nrow = 10, ncol = 10)
+  rast <- raster(mat)
+  v1 <- Moran(rast)
+  v2 <- moranI_matrix_window(mat, window = matrix(1, nrow = 3, ncol = 3))
+  expect_equal(v1, v2)
+})
+
+
+test_that("The local moran I calculated is the same as the one in raster package",{
+  mat <- matrix(sample(1:10, size = 100, replace = TRUE), nrow = 10, ncol = 10)
+  rast <- raster(mat)
+  v1 <- raster::values(MoranLocal(rast))
+  v2 <- local_moranI_matrix_window(mat, window = matrix(1, nrow = 3, ncol = 3))
+  expect_equal(v1, v2)
+})
+
+
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+#### Testing the ELSA function
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+### for vector
+
+test_that("The ELSA index calculated on vector should give the same values as in the original paper",{
+
+  categories <- rep(1, times = 9)
+  categories[[5]] <- 2
+  dists <- rbind(c(0,1), c(1,0))
+  neighmat <- matrix(0, ncol = 9, nrow = 9)
+  p1 <- c(1,1,1,2,2,2,2,2,3,3,3,4,4,4,4,4,5,5,5,5,5,5,5,5,6,6,6,6,6,7,7,7,8,8,8,8,8,9,9,9)
+  p2 <- c(2,4,5,1,4,5,6,3,2,5,6,1,2,5,7,8,1,2,3,4,6,7,8,9,2,3,5,8,9,4,5,8,4,5,6,7,9,5,6,8)
+
[…]

### 2021-07-24 R/boostrap_clust_validation.R
The v2 is almost ready for CRAN !

@@ -59,6 +59,11 @@
 #' }
 boot_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.01, init = "random", verbose = TRUE, seed = NULL){
 
+  if(object$algo %in% c("FCM", "GFCM", "SFCM", "SGFCM") == FALSE){
+    stop('bootstrap group validation can only be performed for FCMres object
+         if algo is one of "FCM", "GFCM", "SFCM", "SGFCM"')
+  }
+
   ## calulating the lagged dataset if necessary -----------------------
   if(object$algo %in% c("SGFCM", "SFCM")){
     wdata <- calcLaggedData(object$Data, object$listw, object$lag_method)
@@ -223,6 +228,11 @@ boot_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.0
 #' if (!inherits(future::plan(), "sequential")) future::plan(future::sequential)
 #' }
 boot_group_validation.mc <- function(object, nsim = 1000, maxiter = 1000, tol = 0.01, init = "random", verbose = TRUE, seed = NULL){
+
+  if(object$algo %in% c("FCM", "GFCM", "SFCM", "SGFCM") == FALSE){
+    stop('bootstrap group validation can only be performed for FCMres object
+         if algo is one of "FCM", "GFCM", "SFCM", "SGFCM"')
+  }
 
   ## calulating the lagged dataset if necessary -----------------------
   if(object$algo %in% c("SGFCM", "SFCM")){

### 2021-07-24 R/data_doc.R

@@ -42,7 +42,6 @@
 #'   \item{near infrared}{Wavelength: 0.85-0.88}
 #'   \item{shortwave infrared}{Wavelength: 1.57-1.65}
 #'   \item{shortwave infrared}{Wavelength: 2.11-2.29}
-#'   ...
 #' }
 #' @source \url{https://earthexplorer.usgs.gov/}
 "Arcachon"

### 2021-07-24 R/side_functions.R

@@ -445,6 +445,7 @@ spiderPlots<- function(data, belongmatrix, chartcolors=NULL){
 #' result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
 #' violinPlots(dataset, result$Groups)
 violinPlots <- function(data,groups){
+    data <- as.data.frame(data)
     data$groups <- groups
     groupvar <- "groups"
     Plots <- list()

### 2021-07-27 R/geocmeans.R
minor corrections in documentation

@@ -13,8 +13,9 @@
 #' alpha = 0, then the spatial c-means is equal to a classical c-means. If alpha
 #' = 1, then the weights given to  X and WX are equals. If alpha = 2, then the
 #' weight of WX is twice the one of X and so on.
-#' An index to measure the spatial consistency of a classification is proposed in
-#' this package.
+#' Several indices are provided to assess the quality of a classification on the
+#' semantic and spatial dimensions. To explore results, a shiny app is also
+#' available
 #'
 #'
 #' @docType package

### 2021-08-08 R/RcppExports.R
Adding non-local spatial function for raster and vector + Minor corrections in boostrap and uncertaintyMap

@@ -154,6 +154,18 @@ focal_euclidean_arr_window <- function(mat, window) {
     .Call(`_geocmeans_focal_euclidean_arr_window`, mat, window)
 }
 
+#' @title focal mean weighted by inverse of euclidean distance on a cube
+#' @name focal_adj_mean_arr_window
+#' @param mat an array (cube)
+#' @param window a numeric matrix (squared)
+#' @return a lagged version of the original cube
+#' @keywords internal
+#' @export
+#'
+focal_adj_mean_arr_window <- function(mat, window) {
+    .Call(`_geocmeans_focal_adj_mean_arr_window`, mat, window)
+}
+
 #' @title Jaccard similarity coefficient
 #' @name calc_jaccard_idx
 #' @description Calculate the Jaccard similarity coefficient

### 2021-08-08 R/boostrap_clust_validation.R

@@ -66,7 +66,19 @@ boot_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.0
 
   ## calulating the lagged dataset if necessary -----------------------
   if(object$algo %in% c("SGFCM", "SFCM")){
-    wdata <- calcLaggedData(object$Data, object$listw, object$lag_method)
+    if(object$isRaster == FALSE){
+      wdata <- calcLaggedData(object$Data, object$listw, object$lag_method)
+    }else{
+      dataset <- lapply(1:ncol(object$Data), function(i){
+        rast <- object$rasters[[1]]
+        vec <- rep(NA, times = raster::ncell(rast))
+        vec[object$missing] <- object$Data[,i]
+        raster::values(rast) <- vec
+        return(rast)
+      })
+      wdata <- calcWdataRaster(object$window, dataset, object$lag_method, object$missing)
+    }
+
   }else{
     wdata <- NULL
   }
@@ -102,20 +114,9 @@ boot_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.0
   }
 
   cons_values <- lapply(all_perm, function(results){
-    # qual_mat <- matrix(-1, nrow = k, ncol = k)
-    # for(ii in 1:k){
-    #   for(jj in 1:k){
-    #     if(qual_mat[ii,jj] == -1){
-    #       x <- results$Belongings[,ii]
-    #       y <- object$Belongings[,jj][results$idx]
-    #       val <- calc_jaccard_idx(x,y)
-    #       qual_mat[ii,jj] <- val
-    #     }
-    #   }
-    # }
     matX <- results$Belongings
     matY <- object$Belongings[results$idx,]
-    qual_mat <- calc_jaccard_mat(matX,matY)
+    qual_mat <- calc_jaccard_mat(matY,matX)
     colnames(qual_mat) <- 1:ncol(qual_mat)
     rownames(qual_mat) <- 1:nrow(qual_mat)
 
@@ -155,20 +156,19 @@ boot_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.0
 
   ## creating a list with the centers values boostraped of clusters -----------------------
   print("Extracting the centers of the clusters...")
-  clust_centers <- lapply(1:k, function(i){
-    idx <- mat_idx[,i]
-    clust_table <- do.call(rbind,lapply(1:nsim, function(j){
-      all_perm[[j]]$Centers[idx[[j]],]
[…]

### 2021-08-08 R/raster_support.R

@@ -16,6 +16,63 @@ check_window <- function(w){
   }
 }
 
+
+#' @title Calculate lagged values for a raster dataset
+#'
+#' @description Calculate lagged values for a raster dataset given a window
+#' and an agregation function
+#'
+#' @param w A matrix
+#' @param dataset A list of rasters
+#' @param fun A string giving the name of a function or a function or "nl"
+#' for non-local method
+#' @param missing_pxl A boolean vector of missing (FALSE) pixels
+#' @keywords internal
+#' @examples
+#' # this is an internal function, no example provided
+calcWdataRaster <- function(w, dataset, fun, missing_pxl){
+
+  if(class(fun) != "function"){
+    fun <- as.character(fun)
+    if(fun != "nl"){
+      useNL <- FALSE
+      tryCatch(fun <- eval(parse(text=fun)),
+               error = function(e)
+                 print("When using rasters, the parameter lag_method must be a function or a string
+                 that can be parsed to a function like sum, mean, min, max ...,
+                 Note that this function is applied to the pixels values multiplied by the weights in the window.
+                 There is one exception : 'nl_mean', see help(Cmeans).")
+      )
+    }else{
+      useNL <- TRUE
+    }
+  }
+  if(useNL == FALSE){
+    Wdatamatrix <- do.call(cbind,lapply(dataset, function(band){
+      wraster <- focal(band, w, fun, na.rm = TRUE, pad = TRUE)
+      return(raster::values(wraster))
+    }))
+  }else{
+    # step1: creating an array with all the matrices
+    mats <- lapply(dataset, raster::as.matrix)
+    arr <- array(do.call(c,mats),c(nrow(mats[[1]]), ncol(mats[[1]]), length(mats)))
+    # step2 : getting the lagged version of the array
+    arr_lag <- focal_adj_mean_arr_window(arr, w)
+    # step3 : creating the Wdatamatrix
+    nsl <- dim(arr_lag)[[3]]
+    Wdatamatrix <- do.call(cbind, lapply(1:nsl, function(i){
+      vec <- arr_lag[,,i]
+      dim(vec) <- NULL
+      return(vec)
+    })
+    )
+  }
+  Wdata_class <- Wdatamatrix[missing_pxl,]
+  colnames(Wdata_class) <- names(dataset)
+  return(Wdata_class)
+}
+
+
 #' @title Check dimensions of a list of rasters
 #'
[…]

### 2021-08-08 man-roxygen/lag_method-arg.R

@@ -3,4 +3,7 @@
 #' can be given (or a string which will be parsed). It will be applied to all the pixels values
 #' in the matrix designated by the parameter window and weighted according to the values of this matrix.
 #' Typically, to obtain an average of the pixels in a 3x3 matrix one could use the function sum (or "sum")
-#' and set the window as: window <- matrix(1/9,nrow = 3, ncol = 3).
+#' and set the window as: window <- matrix(1/9,nrow = 3, ncol = 3). There is one special case when working 
+#' with rasters: one can specify "nl" (standing for non-local) which calculated a lagged version of the 
+#' input rasters, using the inverse of the euclidean distance as spatial weights (see the section Advanced 
+#' examples in the vignette introduction for more details).

### 2021-08-08 src/RcppExports.cpp

@@ -128,6 +128,18 @@ BEGIN_RCPP
     return rcpp_result_gen;
 END_RCPP
 }
+// focal_adj_mean_arr_window
+arma::cube focal_adj_mean_arr_window(arma::cube mat, arma::mat window);
+RcppExport SEXP _geocmeans_focal_adj_mean_arr_window(SEXP matSEXP, SEXP windowSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< arma::cube >::type mat(matSEXP);
+    Rcpp::traits::input_parameter< arma::mat >::type window(windowSEXP);
+    rcpp_result_gen = Rcpp::wrap(focal_adj_mean_arr_window(mat, window));
+    return rcpp_result_gen;
+END_RCPP
+}
 // calc_jaccard_idx
 double calc_jaccard_idx(arma::vec x, arma::vec y);
 RcppExport SEXP _geocmeans_calc_jaccard_idx(SEXP xSEXP, SEXP ySEXP) {
@@ -425,6 +437,7 @@ static const R_CallMethodDef CallEntries[] = {
     {"_geocmeans_focal_euclidean_mat_window", (DL_FUNC) &_geocmeans_focal_euclidean_mat_window, 2},
     {"_geocmeans_focal_euclidean_list", (DL_FUNC) &_geocmeans_focal_euclidean_list, 2},
     {"_geocmeans_focal_euclidean_arr_window", (DL_FUNC) &_geocmeans_focal_euclidean_arr_window, 2},
+    {"_geocmeans_focal_adj_mean_arr_window", (DL_FUNC) &_geocmeans_focal_adj_mean_arr_window, 2},
     {"_geocmeans_calc_jaccard_idx", (DL_FUNC) &_geocmeans_calc_jaccard_idx, 2},
     {"_geocmeans_calc_jaccard_mat", (DL_FUNC) &_geocmeans_calc_jaccard_mat, 2},
     {"_geocmeans_moranI_matrix", (DL_FUNC) &_geocmeans_moranI_matrix, 2},

### 2021-08-21 R/clustering_evaluation.R
many spelling errors corrected after revision

@@ -6,7 +6,7 @@
 #' @description Calculate the Silhouette index of clustering quality.
 #' @details
 #' The index is calculated with the function SIL.F from the package fclust.
-#' When the dataset is too big, an approach by subsampling is used to avoid
+#' When the dataset is too large, an approach by subsampling is used to avoid
 #' crash.
 #' @param data The original dataframe used for the clustering (n*p)
 #' @param belongings A membership matrix (n*k)
@@ -53,7 +53,7 @@ calcSilhouetteIdx <- function(data, belongings){
 #'
 #' @param data The original dataframe used for the clustering (n*p)
 #' @param belongmatrix A membership matrix (n*k)
-#' @param centers The centers of the clusters
+#' @param centers The centres of the clusters
 #' @return A float: the Negentropy Increment index
 #' @export
 #' @examples
@@ -117,7 +117,7 @@ calcNegentropyI <- function(data, belongmatrix, centers){
 #'
 #' @param data The original dataframe used for the clustering (n*p)
 #' @param belongmatrix A membership matrix (n*k)
-#' @param centers The centers of the clusters
+#' @param centers The centres of the clusters
 #' @return A float: the  Generalized Dunn’s index (53)
 #' @export
 #' @examples
@@ -179,7 +179,7 @@ calcGD53 <- function(data, belongmatrix, centers){
 #'
 #' \deqn{\delta_{r}\left(\omega_{i}, \omega_{j}\right)=\left\|\boldsymbol{c}_{i}-\boldsymbol{c}_{j}\right\|}
 #'
-#' which is basically the Euclidean distance between the centers of clusters \eqn{c_{i}} and \eqn{c_{j}}
+#' which is basically the Euclidean distance between the centres of clusters \eqn{c_{i}} and \eqn{c_{j}}
 #'
 #' The denominator is a measure of the maximal dispersion of all clusters, given
 #' by the formula:
@@ -191,7 +191,7 @@ calcGD53 <- function(data, belongmatrix, centers){
 #'
 #' @param data The original dataframe used for the clustering (n*p)
 #' @param belongmatrix A membership matrix (n*k)
-#' @param centers The centers of the clusters
+#' @param centers The centres of the clusters
 #' @return A float: the  Generalized Dunn’s index (43)
 #' @export
 #' @examples
@@ -236,7 +236,7 @@ calcGD43 <- function(data, belongmatrix, centers){
[…]

### 2021-08-21 R/shinyapp.R

@@ -26,8 +26,8 @@ globalVariables(c("spatial4326", "mapfun", "variables", "belongings", "n", "myma
 #' object.
 #' @param dataset A dataframe or matrix representing the data used for the
 #' classification. If NULL, then the matrix is extracted from object.
-#' @param port A integer of length 4 indicating the port on which starting the
-#' shiny app. Default is 8100
+#' @param port An integer of length 4 indicating the port on which to start the
+#' Shiny app. Default is 8100
 #' @param ... Other parameters passed to the function runApp
 #' @importFrom leaflet addRasterImage colorBin leaflet addPolygons addPolylines addCircles addLayersControl hideGroup addLegend addProviderTiles colorFactor
 #' @importFrom grDevices colorRamp

### 2021-08-21 R/spatial_matrices.R

@@ -7,12 +7,12 @@
 #' @details The original function comes from here: https://scrogster.wordpress.com/2012/10/05/applying-a-circular-moving-window-filter-to-raster-data-in-r/
 #' but we reworked it to make it faster and to ensure that the result is a matrix with odd dimensions.
 #'
-#' @param radius The size in meters of the radius of the circular focal
-#' @param res The width in meters of a pixel. It is assumed that pixels are squares.
+#' @param radius The size in metres of the radius of the circular focal
+#' @param res The width in metres of a pixel. It is assumed that pixels are squares.
 #' @return A binary weight matrix
 #' @export
 #' @examples
-#' # wide of 100 meters for pixels of 2 meters
+#' # wide of 100 metres for pixels of 2 metres
 #' window <- circular_window(100, 2)
 #' # row standardisation
 #' window_row_std <- window / sum(window)

### 2021-08-21 man-roxygen/shared_paramsfcm-arg.R

@@ -1,11 +1,11 @@
 #' @param maxiter An integer for the maximum number of iterations
 #' @param tol The tolerance criterion used in the evaluateMatrices function for
 #'   convergence assessment
-#' @param standardize A boolean to specify if the variables must be centered and
+#' @param standardize A boolean to specify if the variables must be centred and
 #'   reduced (default = True)
 #' @param verbose A boolean to specify if the progress should be printed
-#' @param init A string indicating how the initial centers must be selected. "random"
-#' indicates that random observations are used as centers. "kpp" use a distance based method
-#' resulting in more dispersed centers at the beginning. Both of them are heuristic.
+#' @param init A string indicating how the initial centres must be selected. "random"
+#' indicates that random observations are used as centres. "kpp" use a distance-based method
+#' resulting in more dispersed centres at the beginning. Both of them are heuristic.
 #' @param seed An integer used for random number generation. It ensures that the
-#' start centers will be the same if the same value is selected.
+#' starting centres will be the same if the same value is selected.

### 2021-08-21 man-roxygen/window-arg.R

@@ -1,5 +1,5 @@
 #' @param window If data is a list of rasters, then a window must be specified instead of
 #' a list.w object. It will be used to calculate a focal function on each raster. The
-#' window must be a square numeric matrix with odd dimensions (like 3x3). The values in
-#' the matrix indicate the weight to give to each pixel and the center of the matrix is
-#' the center of the focal function.
+#' window must be a square numeric matrix with odd dimensions (such 3x3). The values in
+#' the matrix indicate the weight to give to each pixel and the centre of the matrix is
+#' the centre of the focal function.

### 2021-08-23 README.rmd
many changes and corrections for raster support and extended spatial analysis of results. Part1
Minor edits for CRAN submission !

@@
 
-Philippe Apparcio, Laboratoire d'Équité Environnemental INRS (CANADA), Email: philippe.apparicio@ucs.inrs.ca
+Philippe Apparicio, Laboratoire d'Équité Environnemental INRS (CANADA), Email: philippe.apparicio@ucs.inrs.ca
 
@@
 fuzzy clustering algorithm and to account for the spatial dimension of 
-the dataset. Indexes for measuring the spatial consistency and classification 
-quality are proposed in addition. The algorithms were developed first for 
+the dataset. In addition, indexes for measuring the spatial consistency and classification 
+quality are proposed. The algorithms were developed first for 
 brain imagery as described in the articles of [Cai and al.
@@
 [Gelb and Apparicio](https://doi.org/10.4000/cybergeo.36414) proposed to apply the method to perform a socio-residential and environmental taxonomy in Lyon (France).
+The methods can be applied to dataframes or to rasters.
 
@@
 
+#### Classification consistency
+
+To assess the stability of the obtained clusters, a function for bootstrap validation is proposed: `boot_group_validation`.
+The results can be used to verify if the obtained clusters are stable and how much their centres vary.
+
 #### Interpretation
@@
 
-#### Spatial inconsistency
+#### Spatial diagnostic
+
+Several spatial indices can be calculated to have a better spatial understanding of the obtained clusters, like the global or local Moran I calculated on the membership values, or the join-count-test on the most likely group for each observation.
+ELSA and Fuzzy ELSA statistics can also be calculated to identify areas with high or low multidimensional spatial autocorrelation in the membership values.
+See functions `spConsistency`, `calcELSA`, `calcFuzzyELSA` and `spatialDiag`.
 

### 2021-08-23 R/shinyapp.R

@@ -5,8 +5,15 @@
 # the variables used in the shiny environment must be declared as globalVariables
 globalVariables(c("spatial4326", "mapfun", "variables", "belongings", "n", "mymap",
                   "dataset", "base_violinplots", "dark", "light", "uncertainMap",
-                  "base_boxplots","radarchart", 'rasterMode', "object"
+                  "base_boxplots","radarchart", 'rasterMode', "object","shiny_data"
                   ))
+
+
+#' @title geocmeans general environment
+#'
+#' @description An environment used by geocmeans to store data, functions and values
+#' @keywords internal
+geocmeans_env <- new.env()
 
 #see here to remove global environment variables
 #https://community.rstudio.com/t/alternative-to-global-for-passing-variable-to-shiny-app/26476/2
@@ -50,13 +57,16 @@ globalVariables(c("spatial4326", "mapfun", "variables", "belongings", "n", "myma
 #'
 #' Cmean <- CMeans(Data,4,1.5,500,standardize = FALSE, seed = 456, tol = 0.00001, verbose = FALSE)
 #'
-#' sp_clust_explorer(LyonIris, Cmean)
+#' sp_clust_explorer(Cmean, LyonIris)
 #' }
 sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, dataset = NULL, port = 8100, ...) {
 
   # if(object$isRaster){
   #   stop("The shiny app can not be used currently to display results from raster data, sorry...")
   # }
+
+  # creating a list to store all the data to pass to the shiny app
+  shiny_data <- list()
 
   # checking if the directory of hte shiny app is here  ---------------------------------------
   appDir <- system.file("shiny-examples", "cluster_explorer", package = "geocmeans")
@@ -121,7 +131,8 @@ sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, 
     inertia <- calcexplainedInertia(dataset, belongings)
   }
 
-  assign('inertia', inertia, .GlobalEnv)
+  #assign('inertia', inertia, .GlobalEnv)
+  shiny_data$inertia <- inertia
 
 
   # Preparing some global variables for the app ----------------------------------------
[…]

### 2021-08-23 inst/shiny-examples/cluster_explorer/ui.R

@@ -63,6 +63,14 @@ boxplots_ui <- function(n){
 #' @importFrom leaflet leafletOutput
 #' @importFrom plotly plotlyOutput
 #' @importFrom utils installed.packages
+#'
+
+## extracting all the variables from the shiny_data object
+geocmeans_env <- geocmeans::geocmeans_env
+shiny_data <- geocmeans_env$shiny_data
+for (name in names(shiny_data)){
+  assign(name, shiny_data[[name]])
+}
 
 ## selecting the right number of columns and rows for the violinplots
 nv <- length(variables)

### 2021-10-20 README.rmd
adding codcovr and correcting a minor bug in cat_to_belongings

@@
 ```
+[![Codecov test coverage](https://codecov.io/gh/JeremyGelb/geocmeans/branch/master/graph/badge.svg)](https://codecov.io/gh/JeremyGelb/geocmeans?branch=master)
 <!-- badges: end -->

### 2021-10-20 R/side_functions.R

@@ -34,7 +34,7 @@
 #' result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
 #' MyMaps <- mapClusters(LyonIris, result$Belongings)
 #' }
-mapClusters <- function(geodata = NULL, object, undecided = NULL) {
+mapClusters <- function(geodata = NULL, object, undecided = NULL) {# nocov start
 
     cls <- class(object)[[1]]
     isRaster <- FALSE
@@ -666,7 +666,7 @@ uncertaintyMap <- function(geodata, belongmatrix, njit = 150, radius = NULL, col
 
     return(uncertain_map)
 
-}
+} # nocov end
 
 
 # %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@ -1083,7 +1083,8 @@ adjustSpatialWeights <- function(data,listw,style){
 
 #' @title Convert categories to membership matrix
 #'
-#' @description Function to convert a character vector to a membership matrix (binary matrix)
+#' @description Function to convert a character vector to a membership matrix (binary matrix).
+#' The columns of the matrix are ordered with the order function.
 #'
 #' @param categories A vector with the categories of each observation
 #'
@@ -1091,10 +1092,12 @@ adjustSpatialWeights <- function(data,listw,style){
 #' @export
 cat_to_belongings <- function(categories){
     cats <- unique(categories)
+    cats <- cats[order(cats)]
     cols <- lapply(cats, function(i){
         return (ifelse(categories == i, 1, 0))
     })
     mat <- do.call(cbind, cols)
+    colnames(mat) <- cats
     return(mat)
 }
 

### 2021-10-20 tests/testthat/test_basic_functions.R

@@ -12,3 +12,17 @@ test_that("The euclidean distance function return what is expected",{
   expected <- rep(sum((vec-4)**2), times = 3)
   expect_equal(dists, expected)
 })
+
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+#### Testing the cat_to_belongings
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+test_that("Testing the cat_to_belongings function",{
+  values <- c(1,3,2,1)
+  expected <- cbind(
+    c(1,0,0,1),
+    c(0,0,1,0),
+    c(0,1,0,0)
+  )
+  obtained <- cat_to_belongings(values)
+  expect_true(sum(obtained - expected) == 0)
+})

### 2021-12-22 README.rmd
Adding a pkgdown website
preparing for minor CRAN release

--- 2021-08-23 README.rmd
@@
 ```
+[![Codecov test coverage](https://codecov.io/gh/JeremyGelb/geocmeans/branch/master/graph/badge.svg)](https://app.codecov.io/gh/JeremyGelb/geocmeans?branch=master)
 <!-- badges: end -->
+
+The website of the package is available [here](https://jeremygelb.github.io/geocmeans/)
 

### 2022-02-08 R/FCMclass.R
Correcting bug in issue #2

@@ -205,7 +205,7 @@ is.FCMres <- function(x){
 print.FCMres <- function(x, ...){ #nocov start
   if(x$isRaster){
     part1 <- "FCMres object constructed from a raster dataset"
-    part2 <- paste("dimension of the raster : ",x$rasters[[1]]@nrows,"x",x$rasters[[1]]@cols,
+    part2 <- paste("dimension of the raster : ",x$rasters[[1]]@nrows,"x",x$rasters[[1]]@ncols,
                    " and ", nrow(x$Data), " variables",sep="")
   }else{
     part1 <- "FCMres object constructed from a dataframe"

### 2022-02-08 R/assembling.R

@@ -215,6 +215,9 @@ main_worker <- function(algo, ...){
     centers <- kppCenters(data,k)
   }
 
+  if(is.null(dim(centers))){
+    centers <- matrix(centers, ncol = 1)
+  }
 
   # calculating the first membership matrix
   belongmatrix <- update_belongs(data, centers, dots)

### 2022-10-15 R/FCM.R
Replacing sp by sf, removing maptools and rgeos. Goodby my old friend

@@ -15,7 +15,7 @@
 #' data(LyonIris)
 #' AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img",
 #' "TxChom1564","Pct_brevet","NivVieMed")
-#' dataset <- LyonIris@data[AnalysisFields]
+#' dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' result <- CMeans(dataset,k = 5, m = 1.5, standardize = TRUE)
 CMeans <- function(data, k, m, maxiter = 500, tol = 0.01, standardize = TRUE, verbose = TRUE, init = "random", seed = NULL) {
 
@@ -99,7 +99,7 @@ CMeans <- function(data, k, m, maxiter = 500, tol = 0.01, standardize = TRUE, ve
 #' data(LyonIris)
 #' AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img",
 #' "TxChom1564","Pct_brevet","NivVieMed")
-#' dataset <- LyonIris@data[AnalysisFields]
+#' dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' queen <- spdep::poly2nb(LyonIris,queen=TRUE)
 #' Wqueen <- spdep::nb2listw(queen,style="W")
 #' result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)

### 2022-10-15 R/boostrap_clust_validation.R

@@ -47,7 +47,7 @@
 #'                    "Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")
 #'
 #' #rescaling the columns
-#' Data <- LyonIris@data[AnalysisFields]
+#' Data <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' for (Col in names(Data)){
 #'   Data[[Col]] <- as.numeric(scale(Data[[Col]]))
 #' }
@@ -213,7 +213,7 @@ boot_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.0
 #'                    "Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")
 #'
 #' #rescaling the columns
-#' Data <- LyonIris@data[AnalysisFields]
+#' Data <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' for (Col in names(Data)){
 #'   Data[[Col]] <- as.numeric(scale(Data[[Col]]))
 #' }
@@ -424,7 +424,7 @@ boot_worker <- function(object, wdata, tol, maxiter, init){
 #'                    "Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")
 #'
 #' #rescaling the columns
-#' Data <- LyonIris@data[AnalysisFields]
+#' Data <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' for (Col in names(Data)){
 #'   Data[[Col]] <- as.numeric(scale(Data[[Col]]))
 #' }

### 2022-10-15 R/clustering_evaluation.R

@@ -61,7 +61,7 @@ calcSilhouetteIdx <- function(data, belongings){ #nocov start
 #' data(LyonIris)
 #' AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img",
 #' "TxChom1564","Pct_brevet","NivVieMed")
-#' dataset <- LyonIris@data[AnalysisFields]
+#' dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' queen <- spdep::poly2nb(LyonIris,queen=TRUE)
 #' Wqueen <- spdep::nb2listw(queen,style="W")
 #' result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
@@ -125,7 +125,7 @@ calcNegentropyI <- function(data, belongmatrix, centers){
 #' data(LyonIris)
 #' AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img",
 #' "TxChom1564","Pct_brevet","NivVieMed")
-#' dataset <- LyonIris@data[AnalysisFields]
+#' dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' queen <- spdep::poly2nb(LyonIris,queen=TRUE)
 #' Wqueen <- spdep::nb2listw(queen,style="W")
 #' result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
@@ -199,7 +199,7 @@ calcGD53 <- function(data, belongmatrix, centers){
 #' data(LyonIris)
 #' AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img",
 #' "TxChom1564","Pct_brevet","NivVieMed")
-#' dataset <- LyonIris@data[AnalysisFields]
+#' dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' queen <- spdep::poly2nb(LyonIris,queen=TRUE)
 #' Wqueen <- spdep::nb2listw(queen,style="W")
 #' result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
[…]

### 2022-10-15 R/shinyapp.R

@@ -51,7 +51,7 @@ geocmeans_env <- new.env()
 #'                    "Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")
 #'
 #' #rescaling the columns
-#' Data <- LyonIris@data[AnalysisFields]
+#' Data <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' for (Col in names(Data)){
 #'   Data[[Col]] <- as.numeric(scale(Data[[Col]]))
 #' }
@@ -101,16 +101,16 @@ sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, 
     stop("if object is NULL, spatial must be specified")
   }
 
-  ok_sp <- c("SpatialPolygonsDataFrame", "SpatialPointsDataFrame","SpatialLinesDataFrame")
+  ok_sp <- c("sf")
   if(is.null(object) == FALSE){
     if(object$isRaster == FALSE){
       if(inherits(spatial, ok_sp)  == FALSE){
-        stop('spatial must be one of : c("SpatialPolygonsDataFrame", "SpatialPointsDataFrame","SpatialLinesDataFrame") because object was not created with rasters')
+        stop('spatial must be a feature collection (sf) because object was not created with rasters')
       }
     }
   }else{
     if(inherits(spatial, ok_sp)  == FALSE){
-      stop('spatial must be one of : c("SpatialPolygonsDataFrame", "SpatialPointsDataFrame","SpatialLinesDataFrame")')
+      stop('spatial must be one a feature collection (sf)')
     }
   }
 
@@ -188,13 +188,15 @@ sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, 
     shiny_data$dataset <- dataset
 
     ## creating a referencing raster with the right projection
-    ref_raster <- raster::projectRaster(object$rasters[[1]], crs = sp::CRS("+init=epsg:3857"), method = "ngb")
+    # ref_raster <- raster::projectRaster(object$rasters[[1]], crs = sp::CRS("+init=epsg:3857"), method = "ngb")
+	ref_raster <- raster::projectRaster(object$rasters[[1]], crs = sf::st_crs(3857), method = "ngb")
     #assign('ref_raster', ref_raster, .GlobalEnv)
     shiny_data$ref_raster <- ref_raster
 
     old_names <- names(object$rasters)
[…]

### 2022-10-15 R/side_functions.R

@@ -6,7 +6,7 @@
 #'
 #' @description Build some maps to visualize the results of the clustering
 #'
-#' @param geodata An object of class spatialpolygonesdataframe /
+#' @param geodata An object of class features collection from sf /
 #' spatiallinesdataframe or spatialpointsdataframe ordered
 #' like the original data used for the clustering. Can be Null if object is
 #' a FCMres and has been created with rasters.
@@ -17,9 +17,9 @@
 #'   uncertain (default = NULL)
 #' @return A named list with :
 #' \itemize{
-#'         \item ProbaMaps : a list of ggplot maps showing for each group the
+#'         \item ProbaMaps : a list of tmap maps showing for each group the
 #'         probability of the observations to belong to that group
-#'         \item ClusterMap : a ggplot map showing the most likely group for
+#'         \item ClusterMap : a tmap map showing the most likely group for
 #'          observation
 #' }
 #' @export
@@ -28,7 +28,7 @@
 #' data(LyonIris)
 #' AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img",
 #' "TxChom1564","Pct_brevet","NivVieMed")
-#' dataset <- LyonIris@data[AnalysisFields]
+#' dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' queen <- spdep::poly2nb(LyonIris,queen=TRUE)
 #' Wqueen <- spdep::nb2listw(queen,style="W")
 #' result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
@@ -50,18 +50,31 @@ mapClusters <- function(geodata = NULL, object, undecided = NULL) {# nocov start
     if(isRaster){
         return(mapRasters(object, undecided))
     }else{
-        geodata$OID <- as.character(1:nrow(geodata@data))
-        if(inherits(geodata,"SpatialPolygonsDataFrame")){
-            return(mapPolygons(geodata, belongmatrix, undecided))
-        }else if(inherits(geodata,"SpatialPointsDataFrame")){
-            return(mapPoints(geodata, belongmatrix, undecided))
-        }else if(inherits(geodata,"SpatialLinesDataFrame")){
-            return(mapLines(geodata, belongmatrix, undecided))
-        }else {
+        geodata$OID <- as.character(1:nrow(geodata))
[…]

### 2022-10-15 R/uncertainty.R

@@ -15,7 +15,7 @@
 #' data(LyonIris)
 #' AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img",
 #' "TxChom1564","Pct_brevet","NivVieMed")
-#' dataset <- LyonIris@data[AnalysisFields]
+#' dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' queen <- spdep::poly2nb(LyonIris,queen=TRUE)
 #' Wqueen <- spdep::nb2listw(queen,style="W")
 #' result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
@@ -69,7 +69,7 @@ undecidedUnits <- function(belongmatrix, tol = 0.1, out = "character") {
 #' data(LyonIris)
 #' AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img",
 #' "TxChom1564","Pct_brevet","NivVieMed")
-#' dataset <- LyonIris@data[AnalysisFields]
+#' dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' queen <- spdep::poly2nb(LyonIris,queen=TRUE)
 #' Wqueen <- spdep::nb2listw(queen,style="W")
 #' result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)

### 2022-10-15 tests/testthat/test_belongingmatrices.R

@@ -8,7 +8,7 @@ test_that("Membership matrices must have a rowSums equal to 1",{
   data(LyonIris)
   library(spdep)
   AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")
-  dataset <- LyonIris@data[AnalysisFields]
+  dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
   queen <- poly2nb(LyonIris,queen=TRUE)
   Wqueen <- nb2listw(queen,style="W")
   FCM <- CMeans(dataset, k = 5, m = 1.5, standardize = TRUE)
@@ -31,7 +31,7 @@ test_that("SFCM and FCM should yield identical results if alpha = 0 for SFCM",{
   data(LyonIris)
   library(spdep)
   AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")
-  dataset <- LyonIris@data[AnalysisFields]
+  dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
   queen <- poly2nb(LyonIris,queen=TRUE)
   Wqueen <- nb2listw(queen,style="W")
   result1 <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 0, standardize = TRUE, seed = 123, init = "kpp")
@@ -45,7 +45,7 @@ test_that("GFCM and FCM should yield identical results if beta = 0 for GFCM",{
   data(LyonIris)
   library(spdep)
   AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")
-  dataset <- LyonIris@data[AnalysisFields]
+  dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
   result1 <- GCMeans(dataset, k = 5, m = 1.5, beta = 0, standardize = TRUE, seed = 123)
   result2 <- CMeans(dataset, k = 5, m = 1.5, standardize = TRUE, seed = 123)
   x <- round(result1$Belongings - result2$Belongings,8)
@@ -57,7 +57,7 @@ test_that("SGFCM and FCM should yield identical results if beta = 0 and alpha = 
[…]

### 2022-10-16 README.rmd
Correcting bug in issue #2
New version on cran

@@
 
-# geocmeans <img src='man/figures/geocmeans_logo.png' align="right" height="138.5" />
+# geocmeans <img src='man/figures/geocmeans_logo.png' align="right" style = 'height:138px;'/>
 ## An R package to perform Spatial Fuzzy C-means.
@@
 The website of the package is available [here](https://jeremygelb.github.io/geocmeans/)
+
+## Breaking news
+
+Here we are!
+We are moving from `maptools`, `sp` and `rgeos` to `sf` and `tmap`.
+All the functions and the documentation were modified accordingly.
+If you spot an error or a bug, please open an issue on github.
 

### 2022-10-16 R/clustering_evaluation.R

@@ -608,7 +608,7 @@ spatialDiag <- function(object, nblistw = NULL, window = NULL, undecided = NULL,
       }
     }else{
       if(is.null(object$nblistw) & is.null(nblistw)){
-        stop("The FCMres object was created with a SpatialDataFrame or a dataframe, but does not contain a nblistw slot and the nblistw parameter is NULL. Please specify a nblistw")
+        stop("The FCMres object was created with a feature collection or a dataframe, but does not contain a nblistw slot and the nblistw parameter is NULL. Please specify a nblistw")
       }
     }
     if(is.null(matdist)){

### 2022-10-16 R/side_functions.R

@@ -7,8 +7,7 @@
 #' @description Build some maps to visualize the results of the clustering
 #'
 #' @param geodata An object of class features collection from sf /
-#' spatiallinesdataframe or spatialpointsdataframe ordered
-#' like the original data used for the clustering. Can be Null if object is
+#' ordered like the original data used for the clustering. Can be Null if object is
 #' a FCMres and has been created with rasters.
 #' @param object  A FCMres object, typically obtained from functions CMeans,
 #'   GCMeans, SFCMeans, SGFCMeans. Can also be a simple membership matrix.
@@ -492,7 +491,7 @@ uncertaintyMap <- function(geodata, belongmatrix, njit = 150, radius = NULL, col
             stop("When mapping points, the parameter radius can not be NULL")
         }
     }else{
-        stop("geodata must be one of SpatialPolygonsDataFrame, SpatialLinesDataFrame or SpatialPointsDataFrame")
+        stop("geodata must be a feature collection (POINT, MULTIPOINT, LINESTRING, MULTILINESTRING, POLYGON or MULTIPOLYGON)")
     }
     maxA <- max(geodata$area)
     rt <- as.numeric(njit / maxA)

### 2022-12-18 R/raster_support.R
removing the package raster and rgdal from dependencies

@@ -50,11 +50,11 @@ calcWdataRaster <- function(w, dataset, fun, missing_pxl){
   if(useNL == FALSE){
     Wdatamatrix <- do.call(cbind,lapply(dataset, function(band){
       wraster <- focal(band, w, fun, na.rm = TRUE, pad = TRUE)
-      return(raster::values(wraster))
+      return(terra::values(wraster, mat = FALSE))
     }))
   }else{
     # step1: creating an array with all the matrices
-    mats <- lapply(dataset, raster::as.matrix)
+    mats <- lapply(dataset, terra::as.matrix)
     arr <- array(do.call(c,mats),c(nrow(mats[[1]]), ncol(mats[[1]]), length(mats)))
     # step2 : getting the lagged version of the array
     arr_lag <- focal_adj_mean_arr_window(arr, w)
@@ -83,7 +83,7 @@ calcWdataRaster <- function(w, dataset, fun, missing_pxl){
 #' # this is an internal function, no example provided
 check_raters_dims <- function(rasters){
   dims <- lapply(rasters, function(i){
-    return(c(raster::nrow(i), raster::ncol(i)))
+    return(c(terra::nrow(i), terra::ncol(i)))
   })
   dims <- do.call(rbind, dims)
   ref <- dims[1,]
@@ -104,7 +104,7 @@ check_raters_dims <- function(rasters){
 #' @param standardize A boolean to specify if the variable must be centered and
 #'   reduced (default = True)
 #' @return A list with the required elements to perform clustering
-#' @importFrom raster focal ncell
+#' @importFrom terra focal ncell
 #' @keywords internal
 #' @examples
 #' # this is an internal function, no example provided
@@ -147,13 +147,13 @@ input_raster_data <- function(dataset, w = NULL, fun = sum, standardize = TRUE){
 
   if(standardize){
     for(i in 1:length(dataset)){
-      dataset[[i]] <- raster::scale(dataset[[i]])
+      dataset[[i]] <- terra::scale(dataset[[i]])
     }
   }
 
   #step1 : prepare the regular dataset
   datamatrix <- do.call(cbind,lapply(dataset, function(band){
-    raster::values(band)
+    terra::values(band, mat = FALSE)
   }))
 
   #step2 : only keep the non-missing valuez
@@ -170,11 +170,11 @@ input_raster_data <- function(dataset, w = NULL, fun = sum, standardize = TRUE){
[…]

### 2022-12-18 R/shinyapp.R

@@ -187,15 +187,13 @@ sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, 
     shiny_data$dataset <- dataset
 
     ## creating a referencing raster with the right projection
-    # ref_raster <- raster::projectRaster(object$rasters[[1]], crs = sp::CRS("+init=epsg:3857"), method = "ngb")
-	ref_raster <- raster::projectRaster(object$rasters[[1]], crs = sf::st_crs(3857), method = "ngb")
+	ref_raster <- terra::project(object$rasters[[1]], y = "epsg:3857", method = "near")
     #assign('ref_raster', ref_raster, .GlobalEnv)
     shiny_data$ref_raster <- ref_raster
 
     old_names <- names(object$rasters)
     object$rasters <- lapply(object$rasters, function(rast){
-      #raster::projectRaster(rast, crs = sp::CRS("+init=epsg:3857"), method = "ngb")
-	  raster::projectRaster(rast, crs = sf::st_crs(3857), method = "ngb")
+	  terra::project(object$rasters[[1]],  y = "epsg:3857", method = "near")
     })
     names(object$rasters) <- old_names
 
@@ -323,7 +321,7 @@ sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, 
     ok_names <- name[grepl("group",name, fixed = TRUE)]
     for (name in ok_names){
       rast <- object$rasters[[name]]
-      vals <- raster::values(rast)
+      vals <- terra::values(rast, mat = FALSE)
       pal <- leaflet::colorNumeric(c("#FFFFFF", colors[[i]]),
                                    vals, na.color = "transparent")
       mymap <- mymap %>%
@@ -337,7 +335,7 @@ sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, 
 
     # adding the last layer with the most likely groups
     rast <- object$rasters$Groups
-    vals <- raster::values(rast)
+    vals <- terra::values(rast, mat = FALSE)
     pal <- leaflet::colorNumeric(colors[1:ncol(object$Belongings)],
                                  vals, na.color = "transparent")
     mymap <- mymap %>%
@@ -418,15 +416,15 @@ sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, 
   }else{
     # IF WE ARE IN RASTER MODE
     all_values <- lapply(object$rasters[ok_names], function(rast){
[…]

### 2022-12-18 tests/testthat/test_FCMresclass.R

@@ -97,13 +97,16 @@ test_that("Testing the behavior of the FCMres class object : simple case with ve
 
 test_that("Testing the behavior of the FCMres class object : simple case with raster data",{
 
-  data("Arcachon")
+  library(terra)
+  Arcachon <- terra::rast(system.file("extdata/Littoral4_2154.tif", package = "geocmeans"))
+  names(Arcachon) <- c("blue", "green", "red", "infrared", "SWIR1", "SWIR2")
+
   library(dplyr)
   # loading each raster as a column in a matrix
   # and scale each column
   all_data <- do.call(cbind, lapply(names(Arcachon), function(n){
     rast <- Arcachon[[n]]
-    return(raster::values(raster::scale(rast)))
+    return(terra::values(terra::scale(rast),mat = FALSE))
   }))
 
   # removing the rows with missing values
@@ -116,7 +119,7 @@ test_that("Testing the behavior of the FCMres class object : simple case with ra
   # creating Data (do not forget the standardization)
   Data <- lapply(names(Arcachon), function(n){
     rast <- Arcachon[[n]]
-    return(raster::scale(rast))
+    return(terra::scale(rast))
   })
   names(Data) <- names(Arcachon)
 
@@ -125,12 +128,12 @@ test_that("Testing the behavior of the FCMres class object : simple case with ra
 
   rasters <- lapply(1:7, function(i){
     # creating a vector with only 0 values
-    vals <- rep(0, ncell(ref_raster))
+    vals <- rep(0, terra::ncell(ref_raster))
     # filling it with values when the pixels are not NA
     vals[missing] <- ifelse(kmean7$cluster == i,1,0)
     # setting the values in a rasterLayer
     rast <- ref_raster
-    raster::values(rast) <- vals
+    terra::values(rast) <- vals
     return(rast)
   })
 

### 2022-12-27 R/GFCM.R
Adding the parameters robust and noise_cluster to the main aglrithms. Also, updating the potential JOSS paper. And more documentation.

@@ -19,7 +19,10 @@
 #' "TxChom1564","Pct_brevet","NivVieMed")
 #' dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
 #' result <- GCMeans(dataset,k = 5, m = 1.5, beta = 0.5, standardize = TRUE)
-GCMeans <- function(data, k, m, beta, maxiter = 500, tol = 0.01, standardize = TRUE, verbose = TRUE, init = "random", seed = NULL) {
+GCMeans <- function(data, k, m, beta, maxiter = 500, tol = 0.01, standardize = TRUE,
+                    robust = FALSE,
+                    noise_cluster = FALSE, delta = NULL,
+                    verbose = TRUE, init = "random", seed = NULL) {
 
   #data_class <- class(data)
 
@@ -51,6 +54,8 @@ GCMeans <- function(data, k, m, beta, maxiter = 500, tol = 0.01, standardize = T
                          k = k, m = m, beta = beta,
                          maxiter = maxiter, tol = tol,
                          standardize = standardize, verbose = verbose,
+                         robust = robust,
+                         noise_cluster = noise_cluster, delta = delta,
                          seed = seed, init = init)
 
   # if we are working with rasters, we should set some additional values
@@ -109,8 +114,10 @@ GCMeans <- function(data, k, m, beta, maxiter = 500, tol = 0.01, standardize = T
 #' result <- SGFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, beta = 0.5, standardize = TRUE)
 SGFCMeans <- function(data, nblistw = NULL, k, m, alpha, beta, lag_method="mean",
                       window = NULL,
-                      maxiter = 500, tol = 0.01,
-                      standardize = TRUE, verbose = TRUE, init = "random", seed = NULL) {
+                      maxiter = 500, tol = 0.01, standardize = TRUE,
+                      robust = FALSE,
+                      noise_cluster = FALSE, delta = NULL,
+                      verbose = TRUE, init = "random", seed = NULL) {
 
   isRaster <- inherits(data, "list")
 
@@ -153,6 +160,8 @@ SGFCMeans <- function(data, nblistw = NULL, k, m, alpha, beta, lag_method="mean"
   results <- main_worker('SGFCM', data = data, wdata = wdata,
                          nblistw = nblistw, k = k, m = m, alpha = alpha,
                          beta = beta, lag_method=lag_method, maxiter = maxiter,
+                         robust = robust,
[…]

### 2022-12-27 R/RcppExports.R

@@ -27,12 +27,14 @@ calcCentroids <- function(data, belongmatrix, m) {
 #' @param data A dataframe or matrix representing the observed data with n rows
 #'    and p columns
 #' @param m A float representing the fuzziness degree
+#' @param sigmas A numeric vector for calculating the robust version of the FCM. Filled with ones
+#'    if the classical version is required
 #' @return A n * k matrix representing the probability of belonging of each
 #'    observation to each cluster
 #' @keywords internal
 #'
-calcBelongMatrix <- function(centers, data, m) {
-    .Call(`_geocmeans_calcBelongMatrix`, centers, data, m)
+calcBelongMatrix <- function(centers, data, m, sigmas) {
+    .Call(`_geocmeans_calcBelongMatrix`, centers, data, m, sigmas)
 }
 
 #' @title Calculate the centroids of SFCM
@@ -67,12 +69,60 @@ calcSWFCCentroids <- function(data, wdata, belongmatrix, m, alpha) {
 #' @param alpha A float representing the weight of the space in the analysis (0
 #'   is a typical fuzzy-c-mean algorithm, 1 is balanced between the two
 #'   dimensions, 2 is twice the weight for space)
+#' @param sigmas A numeric vector for calculating the robust version of the FCM. Filled with ones
+#'    if the classical version is required
+#' @param wsigmas Same as sigmas, but calculated on the spatially lagged dataset
 #' @return A n * k matrix representing the belonging probabilities of each
 #'   observation to each cluster
 #' @keywords internal
 #'
-calcSFCMBelongMatrix <- function(centers, data, wdata, m, alpha) {
-    .Call(`_geocmeans_calcSFCMBelongMatrix`, centers, data, wdata, m, alpha)
+calcSFCMBelongMatrix <- function(centers, data, wdata, m, alpha, sigmas, wsigmas) {
+    .Call(`_geocmeans_calcSFCMBelongMatrix`, centers, data, wdata, m, alpha, sigmas, wsigmas)
+}
+
+#' @title Calculate the membership matrix with a noise cluster
+#' @name calcBelongMatrixNoisy
+#' @description Calculate the membership matrix according to a set of centroids, the observed
+#'  data and the fuzziness degree
+#' @param centers A matrix or a dataframe representing the centers of the
+#'    clusters with p columns and k rows
+#' @param data A dataframe or matrix representing the observed data with n rows
+#'    and p columns
+#' @param m A float representing the fuzziness degree
[…]

### 2022-12-27 R/assembling.R

@@ -154,9 +154,21 @@ main_worker <- function(algo, ...){
   k <- dots$k
   update_belongs <- NULL
   udpdate_centers <- NULL
+  robust <- dots$robust
+  dots$sigmas <- rep(1, k)
+  dots$wsigmas <- rep(1, k)
+
+  if(is.null(dots$noise_cluster)){
+    dots$noise_cluster <- FALSE
+  }
 
   #checking if the parameters are ok
   sanity_check(dots,data)
+
+  if(is.null(dots$delta)){
+    dots$delta <- -1
+  }
+
 
   #selecting the functions for calculus
   if(algo == "FCM"){
@@ -204,6 +216,8 @@ main_worker <- function(algo, ...){
     )
   }
 
+  params$robust <- robust
+
   # selecting the original centers from observations
   if (is.null(dots$seed)==FALSE){
     set.seed(dots$seed)
@@ -232,7 +246,18 @@ main_worker <- function(algo, ...){
       setTxtProgressBar(pb, i)
     }
     newcenters <- udpdate_centers(data, centers, belongmatrix, dots)
+
+    # for the robust version, we need to calculate here the sigmas
+    if(robust){
+      dots$sigmas <- calcRobustSigmas(data, belongmatrix, centers, m = dots$m)
+      if(is.null(dots$wdata) == FALSE){
+        dots$wsigmas <- calcRobustSigmas(dots$wdata,belongmatrix, centers, m = dots$m)
+      }
+    }
+
+
     newbelongmatrix <- update_belongs(data, newcenters, dots)
+
     if (evaluateMatrices(belongmatrix, newbelongmatrix, tol) == FALSE) {
       # if we don't reach convergence criterion
       centers <- newcenters
@@ -260,6 +285,11 @@ main_worker <- function(algo, ...){
 
   ## setting the class of the results
   results <- FCMres(results)
+
+  if(dots$noise_cluster){
+    results$noise_cluster <- 1 - rowSums(results$Belongings)
+    results$delta <- dots$delta
+  }
 
   return(results)
 
@@ -323,6 +353,26 @@ sanity_check <- function(dots,data){
     }
   }
 
+  if(!is.null(dots$delta)){
+    if(is.na(dots$delta)){
+      dots$delta <- NULL
+    }
+  }
+
+  # checking delta for clustering with noise
+  if(dots$noise_cluster & is.null(dots$delta)){
[…]

### 2022-12-27 R/data_doc.R

@@ -2,6 +2,7 @@
 #'
 #' @description A dataset containing social and environmental data for the
 #' Iris of Lyon (France)
+#' @name LyonIris
 #'
 #' @format A SpatialPolygonsDataFrame with 506 rows and 32 variables:
 #' \describe{
@@ -34,7 +35,9 @@
 #' The dataset is saved as a Large RasterBrick with the package raster and has the
 #' following crs: EPSG:32630. It is provided as a tiff file.
 #'
-#' @format A Large RasterBrick with 6 bands
+#' @name Arcachon
+#'
+#' @format A spaRast with 6 bands
 #' \describe{
 #'   \item{blue}{wavelength: 0.45-0.51}
 #'   \item{green}{wavelength: 0.53-0.59}
@@ -44,6 +47,19 @@
 #'   \item{shortwave infrared}{wavelength: 2.11-2.29}
 #' }
 #' @source \url{https://earthexplorer.usgs.gov/}
-NULL
+#'
+#' @examples
+#' # loading directly from file
+#' Arcachon <- terra::rast(system.file("extdata/Littoral4_2154.tif", package = "geocmeans"))
+#' names(Arcachon) <- c("blue", "green", "red", "infrared", "SWIR1", "SWIR2")
+#'
+#' # loading with the provided function
+#' Arcachon <- load_arcachon()
+#' @export
+load_arcachon <- function(){
+  Arcachon <- terra::rast(system.file("extdata/Littoral4_2154.tif", package = "geocmeans"))
+  names(Arcachon) <- c("blue", "green", "red", "infrared", "SWIR1", "SWIR2")
+  return(Arcachon)
+}
 
 

### 2022-12-27 R/raster_support.R

@@ -219,9 +219,11 @@ input_raster_data <- function(dataset, w = NULL, fun = sum, standardize = TRUE){
 #' @examples
 #' # this is an internal function, no example provided
 output_raster_data <- function(object, missing, rst){
+
   # object is created by a FCM like function
   # missing is indicating which pixels are complete
   # rst is a basic raster to duplicate it
+  i <- 1
 
   #step1 : creating a raster for each cluster
   rasters <- lapply(1:ncol(object$Belongings), function(i){
@@ -241,6 +243,15 @@ output_raster_data <- function(object, missing, rst){
   vec[missing] <- max.col(DF, ties.method = "first")
   terra::values(rst2) <- vec
   rasters[["Groups"]] <- rst2
+
+  #step3 : adding the noise group if needed
+  if(is.null(object$noise_cluster) == FALSE){
+    rst2 <- rst
+    vec <- rep(NA,times = ncell(rst))
+    vec[missing] <- object$noise_cluster[,i]
+    terra::values(rst2) <- vec
+    object$noise_cluster <- rst2
+  }
 
   object$isRaster <- TRUE
   object$rasters <- rasters

### 2022-12-27 man-roxygen/shared_paramsfcm-arg.R

@@ -3,6 +3,9 @@
 #'   convergence assessment
 #' @param standardize A boolean to specify if the variables must be centred and
 #'   reduced (default = True)
+#' @param robust A boolean indicating if the "robust" version of the algorithm must be used (see details)
+#' @param noise_cluster A boolean indicatong if a noise cluster must be added to the solution (see details)
+#' @param delta A float giving the distance of the noise cluster to each observation
 #' @param verbose A boolean to specify if the progress should be printed
 #' @param init A string indicating how the initial centres must be selected. "random"
 #' indicates that random observations are used as centres. "kpp" use a distance-based method

### 2022-12-27 src/FCM_functions.cpp

@@ -61,18 +61,20 @@ NumericMatrix calcCentroids(arma::mat data, arma::mat belongmatrix, double m){
 //' @param data A dataframe or matrix representing the observed data with n rows
 //'    and p columns
 //' @param m A float representing the fuzziness degree
+//' @param sigmas A numeric vector for calculating the robust version of the FCM. Filled with ones
+//'    if the classical version is required
 //' @return A n * k matrix representing the probability of belonging of each
 //'    observation to each cluster
 //' @keywords internal
 //'
 // [[Rcpp::export]]
-NumericMatrix calcBelongMatrix(NumericMatrix centers, NumericMatrix data, double m){
+NumericMatrix calcBelongMatrix(NumericMatrix centers, NumericMatrix data, double m, NumericVector sigmas){
 
   //calculating euclidean distance between each observation and each center
   NumericMatrix centerdistances(data.nrow(), centers.nrow());
   int i;
   for(i = 0 ; i < centers.nrow() ; i++){
-    centerdistances(_,i) = calcEuclideanDistance2(data, centers(i,_));
+    centerdistances(_,i) = calcEuclideanDistance2(data, centers(i,_)) / sigmas(i);
   }
   NumericMatrix numerator = power_mat(centerdistances, -1.0/(m-1.0));
   NumericVector denom = rowSums(numerator);
@@ -85,7 +87,6 @@ NumericMatrix calcBelongMatrix(NumericMatrix centers, NumericMatrix data, double
   }
   return belongmat;
 }
-
 
 
 //%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@ -144,12 +145,15 @@ NumericMatrix calcSWFCCentroids(arma::mat data, arma::mat wdata, arma::mat belon
 //' @param alpha A float representing the weight of the space in the analysis (0
 //'   is a typical fuzzy-c-mean algorithm, 1 is balanced between the two
 //'   dimensions, 2 is twice the weight for space)
+//' @param sigmas A numeric vector for calculating the robust version of the FCM. Filled with ones
+//'    if the classical version is required
+//' @param wsigmas Same as sigmas, but calculated on the spatially lagged dataset
[…]

### 2022-12-27 src/GFCM_functions.cpp

@@ -17,18 +17,20 @@ using namespace Rcpp;
 //'   and p columns
 //' @param m A float representing the fuzziness degree
 //' @param beta A float for the beta parameter (control speed convergence and classification crispness)
+//' @param sigmas A numeric vector for calculating the robust version of the FCM. Filled with ones
+//'    if the classical version is required
 //' @return A n * k matrix representing the belonging probabilities of each
 //'   observation to each cluster
 //' @keywords internal
 //'
 // [[Rcpp::export]]
-NumericMatrix calcFGCMBelongMatrix(NumericMatrix centers, NumericMatrix data, double m, double beta){
+NumericMatrix calcFGCMBelongMatrix(NumericMatrix centers, NumericMatrix data, double m, double beta, NumericVector sigmas){
 
   //calculating euclidean distance between each observation and each center
   NumericMatrix centerdistances(data.nrow(), centers.nrow());
   int i;
   for(i = 0 ; i < centers.nrow() ; i++){
-    centerdistances(_,i) = calcEuclideanDistance2(data, centers(i,_));
+    centerdistances(_,i) = calcEuclideanDistance2(data, centers(i,_)) / sigmas(i);
   }
 
   //calculating aj
@@ -72,12 +74,15 @@ NumericMatrix calcFGCMBelongMatrix(NumericMatrix centers, NumericMatrix data, do
 //'   is a typical fuzzy-c-mean algorithm, 1 is balanced between the two
 //'   dimensions, 2 is twice the weight for space)
 //' @param beta A float for the beta parameter (control speed convergence and classification crispness)
+//' @param sigmas A numeric vector for calculating the robust version of the FCM. Filled with ones
+//'    if the classical version is required
+//' @param wsigmas Same as sigmas, but calculated on the spatially lagged dataset
 //' @return A n * k matrix representing the belonging probabilities of each
 //'   observation to each cluster
 //' @keywords internal
 //'
 // [[Rcpp::export]]
-NumericMatrix calcSFGCMBelongMatrix(NumericMatrix centers, NumericMatrix data, NumericMatrix wdata, double m, double alpha, double beta){
[…]

### 2022-12-27 src/RcppExports.cpp

@@ -25,15 +25,16 @@ BEGIN_RCPP
 END_RCPP
 }
 // calcBelongMatrix
-NumericMatrix calcBelongMatrix(NumericMatrix centers, NumericMatrix data, double m);
-RcppExport SEXP _geocmeans_calcBelongMatrix(SEXP centersSEXP, SEXP dataSEXP, SEXP mSEXP) {
+NumericMatrix calcBelongMatrix(NumericMatrix centers, NumericMatrix data, double m, NumericVector sigmas);
+RcppExport SEXP _geocmeans_calcBelongMatrix(SEXP centersSEXP, SEXP dataSEXP, SEXP mSEXP, SEXP sigmasSEXP) {
 BEGIN_RCPP
     Rcpp::RObject rcpp_result_gen;
     Rcpp::RNGScope rcpp_rngScope_gen;
     Rcpp::traits::input_parameter< NumericMatrix >::type centers(centersSEXP);
     Rcpp::traits::input_parameter< NumericMatrix >::type data(dataSEXP);
     Rcpp::traits::input_parameter< double >::type m(mSEXP);
-    rcpp_result_gen = Rcpp::wrap(calcBelongMatrix(centers, data, m));
+    Rcpp::traits::input_parameter< NumericVector >::type sigmas(sigmasSEXP);
+    rcpp_result_gen = Rcpp::wrap(calcBelongMatrix(centers, data, m, sigmas));
     return rcpp_result_gen;
 END_RCPP
 }
@@ -53,8 +54,8 @@ BEGIN_RCPP
 END_RCPP
 }
 // calcSFCMBelongMatrix
-NumericMatrix calcSFCMBelongMatrix(NumericMatrix centers, NumericMatrix data, NumericMatrix wdata, double m, double alpha);
-RcppExport SEXP _geocmeans_calcSFCMBelongMatrix(SEXP centersSEXP, SEXP dataSEXP, SEXP wdataSEXP, SEXP mSEXP, SEXP alphaSEXP) {
+NumericMatrix calcSFCMBelongMatrix(NumericMatrix centers, NumericMatrix data, NumericMatrix wdata, double m, double alpha, NumericVector sigmas, NumericVector wsigmas);
[…]

### 2022-12-27 src/shared_functions.cpp

@@ -203,7 +203,6 @@ NumericMatrix pow_matrix_bycol(NumericMatrix x, float p){
 
 
 
-
 //' @title minimum of each row of a matrix
 //' @name rowmins_mat
 //' @param x a matrix
@@ -278,3 +277,39 @@ NumericMatrix vector_out_prod(NumericVector x){
   }
   return res;
 }
+
+// ############################################################################
+// FUNCTIONS FOR ROBUST CASES
+// ############################################################################
+
+//' @title Calculate sigmas for the robust version of the c-means algorithm
+//' @name calcRobustSigmas
+//' @description Calculate sigmas for the robust version of the c-means algorithm
+//' @param data A Numeric matrix representing the observed data with n rows
+//'   and p columns
+//' @param belongmatrix A n X k matrix giving for each observation n, its
+//'   probability to belong to the cluster k
+//' @param centers A c X k matrix giving for each cluster c, its center in k dimensions
+//' @param m A float representing the fuzziness degree
+//' @return A vector with the sigmas for each cluster
+//' @keywords internal
+//
+// [[Rcpp::export]]
+NumericVector calcRobustSigmas(NumericMatrix data, NumericMatrix belongmatrix, NumericMatrix centers, double m){
+
+  NumericMatrix powered = power_mat(belongmatrix, m);
+
+  int nc = centers.nrow();
+  NumericVector sigmas(nc);
+  int i;
+
+  for( i = 0; i < nc ; i++){
+    double denom = sum(powered(_,i));
+    double num = sum(calcEuclideanDistance2(data, centers(i,_)) * powered(_,i));
+    sigmas(i) = num / denom;
+  };
+
+  return(sqrt(sigmas));
+}
+
+

### 2022-12-27 src/shared_functions.h

@@ -31,10 +31,14 @@ LogicalMatrix test_inferior_mat(NumericMatrix mat, double t);
 
 NumericMatrix vector_out_prod(NumericVector x);
 
+double calcDeltaNoiseClust(NumericMatrix data, NumericMatrix centers, double lambda);
+
 double max_mat(NumericMatrix x);
 
 double vecmin(NumericVector x);
 
 double vecmax(NumericVector x);
+
+NumericVector calcRobustSigmas(NumericMatrix data, NumericMatrix belongmatrix, NumericMatrix centers, double m);
 
 #endif

### 2022-12-27 tests/testthat/test_robust_fcm.R

@@ -1,0 +1,40 @@
+context("robust and noisy FCM tests")
+
+
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+#### Testing the function producing the sigmas
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+test_that("The calcRobustSigmas function must returned what is expected",{
+
+ data <- rbind(
+   c(1,2,3),
+   c(1,1,1),
+   c(1,2,2),
+   c(1,1,1),
+   c(1,2,3)
+ )
+ belong_matrix <- rbind(
+   c(0.2,0.8),
+   c(0.4,0.6),
+   c(0.5,0.5),
+   c(0.5,0.5),
+   c(0.7,0.3)
+ )
+ m <- 1.5
+
+ centers <- calcCentroids(data, belong_matrix, m)
+ powered <- belong_matrix**m
+
+ expected <- sqrt(sapply(1:2, function(i){
+   ui <- powered[,i]
+   denom <- sum(ui)
+   num <- sum(calcEuclideanDistance2(data, centers[i,]) * ui)
+   return(num/denom)
+ }))
+
+ obtained <- calcRobustSigmas(data, belong_matrix, centers, m)
+
+ expect_equal(expected, obtained)
+
+})

### 2023-01-29 README.rmd
removing the package raster and rgdal from dependencies
updating the readme

@@
 Here we are!
-We are moving from `maptools`, `sp` and `rgeos` to `sf` and `tmap`.
+We are moving from `maptools`, `sp`, `rgeos`, `raster` and `rgdal` to `sf`, `terra` and `tmap`.
 All the functions and the documentation were modified accordingly.
@@
 Each function return a membership matrix, the data used for the classification (scaled if required) and the centers of the clusters.
+
+For each algorithm, it is possible to calculate a "robust version" and to add a noise group (used to catch outliers).
+See the parameters `robust` and `noise_cluser` in the documentation for more details.
 
@@
 
-`geocmeans` version 0.1.0 is licensed under [GPL2 License](https://github.com/JeremyGelb/geocmeans/blob/master/LICENSE.txt).
+`geocmeans` version 0.3.2 is licensed under [GPL2 License](https://github.com/JeremyGelb/geocmeans/blob/master/LICENSE.txt).

### 2023-01-29 R/raster_support.R
correcting a bug in the shiny app

@@ -213,7 +213,7 @@ output_raster_data <- function(object, missing, rst){
   if(is.null(object$noise_cluster) == FALSE){
     rst2 <- rst
     vec <- rep(NA,times = ncell(rst))
-    vec[missing] <- object$noise_cluster[,i]
+    vec[missing] <- object$noise_cluster
     terra::values(rst2) <- vec
     object$noise_cluster <- rst2
   }

### 2023-01-29 R/shinyapp.R

@@ -134,6 +134,19 @@ sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, 
   #assign('inertia', inertia, .GlobalEnv)
   shiny_data$inertia <- inertia
 
+  # # Adjusting the membership matrix if we have a noise cluster ----------------------------------------
+  noise_mode <- FALSE
+  if(is.null(object) == FALSE){
+    if(is.null(object$noise_cluster) == FALSE){
+      #test <- matrixStats::rowMaxs(object$Belongings) < object$noise_cluster
+      #object$Belongings <- cbind(object$Belongings, object$noise_cluster)
+      #object$Groups[test] <- "noise"
+      noise_vec <- object$noise_cluster
+      noise_mode <- TRUE
+    }
+  }
+
+
 
   # Preparing some global variables for the app ----------------------------------------
   if(is.matrix(dataset)){
@@ -173,6 +186,12 @@ sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, 
   if(is.null(object) == FALSE){
     if(object$isRaster){
       rasterMode <- TRUE
+
+      # let me check if the installed version of leaflet is enough
+      if(packageVersion("leaflet") <= "2.1.1"){
+        stop("To use the shiny app with raster data, you must have installed a version of leaflet > 2.1.1 (ex 2.1.1.9000 from github)")
+      }
+
     }
   }
 
@@ -187,13 +206,13 @@ sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, 
     shiny_data$dataset <- dataset
 
     ## creating a referencing raster with the right projection
-	ref_raster <- terra::project(object$rasters[[1]], y = "epsg:3857", method = "near")
+	  ref_raster <- terra::project(object$rasters[[1]], y = "epsg:3857", method = "near")
     #assign('ref_raster', ref_raster, .GlobalEnv)
     shiny_data$ref_raster <- ref_raster
 
     old_names <- names(object$rasters)
     object$rasters <- lapply(object$rasters, function(rast){
-	  terra::project(object$rasters[[1]],  y = "epsg:3857", method = "near")
+	  terra::project(rast,  y = "epsg:3857", method = "near")
     })
     names(object$rasters) <- old_names
 
@@ -234,7 +253,12 @@ sp_clust_explorer <- function(object = NULL, spatial = NULL, membership = NULL, 
[…]

### 2023-08-23 README.rmd
small changes in documentation

@@
 
+#### Reproductibility
+
+Clustering methods like CMeans depend on the initial centers selected.
+In `geocmeans`, they are selected randomly, and two runs of the functions can yield different results.
+To facilitate the reproductibility of the results, the main functions of the package (`CMeans`, `GFCMeans`, `SFCMeans`, `SGFCMeans`, `selectParameters`, `selectParameters.mc`) have a `seed` parameter.
+It can be set by the user to ensure that the results of the functions are exactly the same.
+
 #### Interpretation

### 2023-08-23 R/RcppExports.R

@@ -347,7 +347,7 @@ Elsa_fuzzy_matrix_window <- function(mats, window, dist) {
 #' @name adj_spconsist_arr_window_globstd
 #' @param data an arma cube of dimension nr,nc,ns
 #' @param memberships an arma cube of dimension nr, nc, ks
-#' @param a matrix representing the neighbouring of each pixel
+#' @param window a matrix representing the neighbouring of each pixel
 #' @param mindist A minimum value for distance between two observations. If two
 #'   neighbours have exactly the same values, then the euclidean distance
 #'   between them is 0, leading to an infinite spatial weight. In that case,

### 2023-09-12 vignettes/FCMres.R
removing vignettes from CRAN release. They will be only available on the website now

@@ -1,156 +1,0 @@
-## ----setup, include=FALSE-----------------------------------------------------
-knitr::opts_chunk$set(echo = TRUE)
-
-## ----message=FALSE, warning=FALSE---------------------------------------------
-library(geocmeans)
-library(tmap)
-library(dplyr)
-library(ggplot2)
-library(spdep)
-library(raster)
-
-data("LyonIris")
-
-# selecting the columns for the analysis
-AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14",
-                   "Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")
-
-# rescaling the columns
-Data <- LyonIris@data[AnalysisFields]
-for (Col in names(Data)){
-  Data[[Col]] <- scale(Data[[Col]])
-}
-
-# applying the hclust function
-clust <- hclust(dist(Data), method = "ward")
-
-# getting the groups
-LyonIris$Hclust_groups <- as.character(cutree(clust, k = 4))
-Data$Hclust_groups <- as.character(cutree(clust, k = 4))
-
-# mapping the groups
-tm_shape(LyonIris) + 
-  tm_polygons(col = "Hclust_groups", title = "groups")
-
-
-## ----message=FALSE, warning=FALSE---------------------------------------------
-centers <- Data %>% 
-  group_by(Hclust_groups) %>%
-  summarise_all(mean)
-
-centers <- as.matrix(centers[2:ncol(centers)])
-
-## ----message=FALSE, warning=FALSE---------------------------------------------
-member_mat <- cat_to_belongings(Data$Hclust_groups)
-
-## ----message=FALSE, warning=FALSE---------------------------------------------
-Data$Hclust_groups <- NULL
-
-hclustres <- FCMres(list(
-  "Centers" = centers,
-  "Belongings" = member_mat,
-  "Data" = Data,
-  "m" = 1,
-  "algo" = "hclust"
-))
-
-## ----message=FALSE, warning=FALSE, eval = FALSE-------------------------------
-#  # quick summaries about the groups
-#  summary(hclustres)
-#  violinPlots(hclustres$Data, hclustres$Groups)
-#  spiderPlots(hclustres$Data, hclustres$Belongings)
-#  mapClusters(LyonIris, hclustres)
-#  
-#  # some indices about classification quality
-#  calcqualityIndexes(hclustres$Data,
[…]

### 2023-09-12 vignettes/introduction.R

@@ -1,343 +1,0 @@
-## ----setup, include=FALSE-----------------------------------------------------
-knitr::opts_chunk$set(echo = TRUE)
-
-## ----message=FALSE, warning=FALSE---------------------------------------------
-#charging packages and data
-library(geocmeans)
-library(ggplot2)
-library(ggpubr)
-library(dplyr)
-library(viridis)
-library(spdep)
-
-data(LyonIris)
-
-# selecting the columns for the analysis
-AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14",
-                   "Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")
-
-# rescaling the columns
-Data <- LyonIris@data[AnalysisFields]
-for (Col in names(Data)){
-  Data[[Col]] <- scale(Data[[Col]])
-}
-
-# preparing some elements for further mapping
-LyonIris$OID <- as.character(1:nrow(LyonIris))
-FortiData <- ggplot2::fortify(LyonIris,region="OID")
-
-## ----include=FALSE------------------------------------------------------------
-# loading the pre-calculated results
-load(system.file("extdata", "results_vignette_intro.rda",
-                           package = "geocmeans", mustWork = TRUE))
-
-## ----warning=FALSE, fig.cap = "Impact of the number of groups on the explained variance", out.width = "50%", fig.pos="H", fig.align="center"----
-# finding the best k by using the r2 of the classification
-# trying for k from 2 to 10
-R2s <- sapply(2:10,function(k){
-  Clust <- kmeans(Data,centers=k,iter.max = 150)
-  R2 <- Clust$betweenss / Clust$totss
-  return(R2)
-})
-
-Df <- data.frame(K=2:10,
-                 R2 = R2s)
-
-ggplot(Df)+
-  geom_line(aes(x=K,y=R2s))+
-  geom_point(aes(x=K,y=R2s),color="red")+
-  xlab("Number of groups")+
-  ylab("R2 of classification")
-
-## ----warning=FALSE, fig.cap="Clusters identified by classical k-means", fig.pos="H", fig.align="center"----
-KMeanClust <-  kmeans(Data,centers=4,iter.max = 150)
-LyonIris$Cluster <-paste("cluster",KMeanClust$cluster,sep="_")
-
-# mapping the groups
[…]

### 2023-09-12 vignettes/rasters.R

@@ -1,57 +1,0 @@
-## ----setup, include=FALSE-----------------------------------------------------
-knitr::opts_chunk$set(echo = TRUE)
-
-## ----include=FALSE------------------------------------------------------------
-#loading the pre-calculated results
-load(system.file("extdata", "results_vignette_raster.rda",
-                           package = "geocmeans", mustWork = TRUE))
-
-## ----message=FALSE, warning=FALSE, fig.width= 4-------------------------------
-library(raster)
-library(geocmeans)
-library(ggpubr)
-library(future)
-library(tmap)
-library(viridis)
-library(RColorBrewer)
-
-data("Arcachon")
-
-# show the pseudo-color image
-plotRGB(Arcachon, r = 3, g = 2, b = 1, stretch = "hist")
-
-## ----message=FALSE, warning=FALSE---------------------------------------------
-# sonverting the RasterBrick to a simple list of RasterLayer
-dataset <- lapply(names(Arcachon), function(n){
-  aband <- Arcachon[[n]]
-  return(aband)
-})
-
-# giving a name to each band
-names(dataset) <- names(Arcachon)
-
-## ----message=FALSE, warning=FALSE, eval=FALSE---------------------------------
-#  # finding an appropriate k and m values (using a multicore plan)
-#  future::plan(future::multisession(workers = 6))
-#  FCMvalues <- select_parameters.mc(algo = "FCM", data = dataset,
-#                                 k = 5:10, m = seq(1.1,2,0.1), spconsist = FALSE,
-#                                 indices = c("XieBeni.index", "Explained.inertia",
-#                                             "Negentropy.index", "Silhouette.index"),
-#                                 verbose = TRUE)
-
-## ----message=FALSE, warning=FALSE, fig.width = 5, fig.align='center'----------
-# plotting the silhouette index values
-ggplot(FCMvalues) + 
-  geom_raster(aes(x = m, y = k, fill = Silhouette.index)) + 
-  geom_text(aes(x = m, y = k, label = round(Silhouette.index,2)), size = 2)+
-  scale_fill_viridis() +
-  coord_fixed(ratio=0.125) 
-
-# plotting the explained inertia
-ggplot(FCMvalues) + 
-  geom_raster(aes(x = m, y = k, fill = Explained.inertia)) + 
-  geom_text(aes(x = m, y = k, label = round(Explained.inertia,2)), size = 2)+
-  scale_fill_viridis() +
-  coord_fixed(ratio=0.125)
-
-

### 2023-09-22 README.rmd
removing vignettes from CRAN release. They will be only available on the website now
edit for publication in JOSS

@@
 [![Codecov test coverage](https://codecov.io/gh/JeremyGelb/geocmeans/branch/master/graph/badge.svg)](https://app.codecov.io/gh/JeremyGelb/geocmeans?branch=master)
+[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.8316593.svg)](https://doi.org/10.5281/zenodo.8316593)
+
 <!-- badges: end -->
@@
 
-`geocmeans` version 0.3.2 is licensed under [GPL2 License](https://github.com/JeremyGelb/geocmeans/blob/master/LICENSE.txt).
+`geocmeans` version 0.3.4 is licensed under [GPL2 License](https://github.com/JeremyGelb/geocmeans/blob/master/LICENSE.txt).

### 2024-11-17 R/boostrap_clust_validation.R
correcting bug from issue 8

@@ -74,7 +74,7 @@ boot_group_validation <- function(object, nsim = 1000, maxiter = 1000, tol = 0.0
         rast <- object$rasters[[1]]
         vec <- rep(NA, times = terra::ncell(rast))
         vec[object$missing] <- object$Data[,i]
-        terra::values(rast, mat = FALSE) <- vec
+        terra::values(rast) <- vec
         return(rast)
       })
       wdata <- calcWdataRaster(object$window, dataset, object$lag_method, object$missing)
@@ -244,7 +244,7 @@ boot_group_validation.mc <- function(object, nsim = 1000, maxiter = 1000, tol = 
         rast <- object$rasters[[1]]
         vec <- rep(NA, times = terra::ncell(rast))
         vec[object$missing] <- object$Data[,i]
-        terra::values(rast, mat = FALSE) <- vec
+        terra::values(rast) <- vec
         return(rast)
       })
       wdata <- calcWdataRaster(object$window, dataset, object$lag_method, object$missing)

### 2024-11-17 R/compositional_methods.R

@@ -1,0 +1,236 @@
+# # data(LyonIris)
+# # AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14","Pct_65","Pct_Img",
+# #                    "TxChom1564","Pct_brevet","NivVieMed")
+# # dataset <- sf::st_drop_geometry(LyonIris[AnalysisFields])
+# # queen <- spdep::poly2nb(LyonIris,queen=TRUE)
+# # Wqueen <- spdep::nb2listw(queen,style="W")
+# # result <- SFCMeans(dataset, Wqueen,k = 5, m = 1.5, alpha = 1.5, standardize = TRUE)
+# #
+# # library(compositions)
+# #
+# # x <- result$Belongings
+# # listw <- Wqueen
+# #
+# #
+# #
+# # cmp_geary <- function(x, listw){
+# #
+# #   cmp <- acomp(x)
+# #   x_clr <- as.matrix(clr(cmp))
+# #
+# #   sums_ij <- sapply(1:nrow(x_clr), function(i){
+# #     ri <- x_clr[i,]
+# #     diffs <- sapply(listw$neighbours[[i]], function(j){
+# #       rj <- x_clr[j,]
+# #       sum((ri - rj) ** 2)
+# #     })
+# #     sub_sum <- sum(listw$weights[[i]] * diffs)
+# #     return(sub_sum)
+# #   })
+# #
+# #   numerator <- (nrow(x_clr) - 1) * sum(sums_ij)
+# #   S0 <- sum(sapply(listw$weights, sum))
+# #   x_bar <- colMeans(x_clr)
+# #   diff_mean <-  sum(c(sweep(x_clr, 2, x_bar, "-")**2))
+# #
+# #   C <- numerator / (2*S0 * diff_mean)
+# #
+# # }
+#
+#
+# spConsistency <- function(object, nblistw = NULL, window = NULL, nrep = 999, adj = FALSE, mindist = 1e-11, use_clr = FALSE) {
+#
+#   if(inherits(object, "FCMres")){
+#     belongmat <- as.matrix(object$Belongings)
+#     if(object$isRaster & is.null(window)){
+#       window <- object$window
+#       if(is.null(window)){
+#         stop("impossible to find a window in the given object, please
+#              specify one by hand.")
+#       }
+#     }
+#     if(object$isRaster == FALSE & is.null(nblistw)){
+#       nblistw <- object$nblistw
+#     }
+#   }else{
[…]

### 2024-11-17 R/spatial_indices.R

@@ -87,6 +87,13 @@ spConsistency <- function(object, nblistw = NULL, window = NULL, nrep = 999, adj
     weights <- nblistw$weights
     neighbours <- nblistw$neighbours
     ## calcul de l'inconsistence spatiale actuelle
+
+    # we could aslo use the Aitchison distance (https://ima.udg.edu/~barcelo/index_archivos/Measures_of_difference__Clustering.pdf)
+    # this is simply done by calculating the clr transformation on the original data
+    # belongmat <- log(belongmat)
+    # belongmat <- sweep(belongmat, 1, rowMeans(belongmat), "-")
+    # belongmat <- as.matrix(belongmat)
+
     obsdev <- sapply(1:nrow(belongmat), function(i) {
       row <- belongmat[i, ]
       idneighbour <- neighbours[[i]]
@@ -95,6 +102,7 @@ spConsistency <- function(object, nblistw = NULL, window = NULL, nrep = 999, adj
         neighbour <- t(as.matrix(neighbour))
       }
       W <- weights[[i]]
+      # we are using here the euclidean distance
       diff <- (neighbour-row[col(neighbour)])**2 * W
       tot <- sum(rowSums(diff))
       return(tot)
@@ -133,6 +141,9 @@ spConsistency <- function(object, nblistw = NULL, window = NULL, nrep = 999, adj
     rasters <- object$rasters[ok_names]
     matrices <- lapply(rasters, terra::as.matrix, wide = TRUE)
     mat_dim <- dim(matrices[[1]])
+
+
+    # applying the
 
     if(adj){
       dataset <- lapply(1:ncol(object$Data), function(ic){
